\chapter{Rings}
\section{Basic Definition}


The concept of a ring generalizes the arithmetic properties we observe in the integers, allowing us to study structures with both addition and multiplication.

\begin{definition}[Ring]
A \textbf{ring} $(R, +, \cdot)$ is a set $R$ equipped with two binary operations, addition ($+$) and multiplication ($\cdot$), such that:
\begin{enumerate}
    \item $(R, +)$ is an \textbf{abelian group} with additive identity $0_R$.
    \item Multiplication is \textbf{associative}: $a \cdot (b \cdot c) = (a \cdot b) \cdot c$ for all $a, b, c \in R$.
    \item The \textbf{distributive laws} hold for all $a, b, c \in R$:
    \[ a \cdot (b + c) = a \cdot b + a \cdot c \quad \text{and} \quad (a + b) \cdot c = a \cdot c + b \cdot c \]
\end{enumerate}
\end{definition}

\begin{example}
\begin{itemize}
    \item The sets $\mathbb{Z}, \mathbb{Q}, \mathbb{R},$ and $\mathbb{C}$ with standard addition and multiplication are rings.
    \item \textbf{Gaussian Integers:} $\mathbb{Z}[i] = \{ a + bi \mid a, b \in \mathbb{Z} \}$. Addition and multiplication follow the rules of complex numbers.
    \item \textbf{Cyclotomic Integers:} Let $\zeta_n = e^{2\pi i/n}$. Then $\mathbb{Z}[\zeta_n] = \{ a_0 + a_1\zeta_n + \dots + a_{n-1}\zeta_n^{n-1} \mid a_j \in \mathbb{Z} \}$ is a ring. For $n=4$, $\zeta_4 = i$, we recover the Gaussian integers.
    \item \textbf{Polynomial Rings:} $\mathbb{Z}[x]$ is the set of all polynomials with integer coefficients. Multiplication is defined by the Cauchy product:
    \[ \left( \sum a_p x^p \right) \left( \sum b_q x^q \right) = \sum_{i} \left( \sum_{p+q=i} a_p b_q \right) x^i \]
    \item \textbf{Matrix Rings:} $M_{n \times n}(\mathbb{Z})$, the set of $n \times n$ matrices with integer entries, is a ring under matrix addition and multiplication.
    \item \textbf{Multiple Integers:} $n\mathbb{Z} = \{ nk \mid k \in \mathbb{Z} \}$ is a ring for any $n \in \mathbb{Z}$.
    \item \textbf{Non-example:} The set of odd integers $\{\dots, -3, -1, 1, 3, \dots\}$ is \textbf{not} a ring because it is not closed under addition (the sum of two odd integers is even).
\end{itemize}
\end{example}

\begin{remark}[Motivation]
The study of rings like $\mathbb{Z}[i]$ or $\mathbb{Z}[\zeta_n]$ is motivated by number theory.
\begin{itemize}
    \item To find integer solutions to $x^2 + y^2 = z^2$, we can factor the left side in $\mathbb{Z}[i]$ as $(x+iy)(x-iy) = z^2$. This leads to the question: does the \textbf{Fundamental Theorem of Arithmetic} hold in $\mathbb{Z}[i]$? (i.e., is there unique factorization into primes?)
    \item In Fermat's Last Theorem, we want to study integer solutions of 
    $$x^n + y^n = z^n,$$ 
    In the ring $\mathbb{Z}[\zeta_n]$, one can factor the sum as $x^n + y^n = \prod_{k=0}^{n-1} (x + \zeta_n^k y)$.
\end{itemize}
\end{remark}


While all rings satisfy the basic axioms, we often focus on rings with additional properties.
\begin{definition}[Unital and Commutative Rings]
Let $R$ be a ring.
\begin{enumerate}
    \item $R$ is \textbf{unital} if there exists a multiplicative identity $1_R \in R$ such that $1_R \cdot r = r \cdot 1_R = r$ for all $r \in R$.
    \item If $R$ is unital, an element $u \in R$ is called a \textbf{unit} if it has a multiplicative inverse; that is, there exists $u^{-1} \in R$ such that $u u^{-1} = u^{-1} u = 1_R$. The set of all units is denoted $U(R)$.
    \item $R$ is \textbf{commutative} if $a \cdot b = b \cdot a$ for all $a, b \in R$.
\end{enumerate}
\end{definition}

\begin{example}
\begin{itemize}
    \item $\mathbb{Z}_n$ is a commutative unital ring. Its units are $U(\mathbb{Z}_n) = \{ [a] \in \mathbb{Z}_n \mid \gcd(a, n) = 1 \}$.
    \item $M_{n \times n}(\mathbb{R})$ is unital but \textbf{not} commutative for $n \ge 2$. Its units are $U(M_{n \times n}(\mathbb{R})) = GL(n, \mathbb{R})$.
    \item $(\mathbb{Z}_n,+,\cdot)$ is a commutative and unital ring with the multiplication identity $1_R=[1]$. The units of $\mathbb{Z}_n$ are $U(\mathbb{Z})=\mathbb{Z}_n^*=\{a\mid gcd(a,n)=1\}$
    \item $R = \mathbb{Z}$ is commutative unital, with $U(R) = \{ \pm 1 \}$.
    \item $R = \mathbb{Z}[i]$ is commutative unital, with $U(R) = \{ \pm 1, \pm i \}$.
    \item For $R =\mathbb{Q}$, $\mathbb{R}$, $\mathbb{C}$, all nonzero elements are units.
    \item $n\mathbb{Z}$ for $n > 1$ is a commutative ring that is \textbf{not} unital.
\end{itemize}
\end{example}


\begin{remark}
\begin{itemize}
    \item The additive identity $0_R \in R$ is unique in $R$. If $R$ is unital, then the multiplication identity $1_R\in R$ is also unique.
    \item We write $-r\in R $ to be the additive inverse of $r\in R$, i.e.: $(-r)+r=r+(-r)=0_R$
    \item If $r\in U(R)$ is a unit in a (commutative) unital ring, then we write $r^{-1}$ to be the multiplication inverse of $r$, i.e.: $r^{-1}r=rr^{-1}=1_R$
    \item For $n\in \mathbb{N}$ and $r\in R$, we write $n\cdot r:=r+\cdots + r$, and $-n\cdot r := (-r) + \dots + (-r)$ ($n$ terms in both sums).
    \item For $a,b\in R$, we write 
$$a\mid b$$ if there exists $c\in R$ such that $ac=b$.
\end{itemize}
\end{remark}

\begin{proposition}
In any unital ring $R$:
\begin{enumerate}
    \item $0_R \cdot r = r \cdot 0_R = 0_R$ for all $r \in R$.
    \item $(-1_R) \cdot r = -r$ (if $R$ is unital).
    \item $(-r) \cdot (-s) = r \cdot s$.
\end{enumerate}
\end{proposition}

\begin{proof}
(1) Since $0_R = 0_R + 0_R$, we have $0_R \cdot r = (0_R + 0_R) \cdot r = 0_R \cdot r + 0_R \cdot r$. By the cancellation law in the additive group $(R, +)$, we obtain $0_R \cdot r = 0_R$.

(2) $0_R = (1_R + (-1_R)) \cdot r = 1_R \cdot r + (-1_R) \cdot r = r + (-1_R) \cdot r$. Thus, $(-1_R) \cdot r$ is the additive inverse of $r$, so $(-1_R) \cdot r = -r$. 

(3) Exercise.
\end{proof}

\begin{definition}[Subring]
A subset $S \leq R$ is a \textbf{subring} if $S$ itself is a ring under the operations of $R$.
\end{definition}

\begin{proposition}[Subring Test]
A non-empty subset $S \subseteq R$ is a subring if and only if for all $a, b \in S$:
\begin{enumerate}
    \item $a - b \in S$
    \item $a \cdot b \in S$
\end{enumerate}
\end{proposition}


\begin{example}
\begin{itemize}
    \item $\mathbb{Z}[i]\leq \mathbb{C}$ is a subring.
    \item $n\mathbb{Z}\leq \mathbb{Z}$ is a subring.
    \item $\{\cdots,-3,-1,1,3,\cdots\}\subset\mathbb{Z}$ is \textbf{NOT} a subring.
\end{itemize}
\end{example}


\begin{definition}[Field]
A \textbf{field} $F$ is a commutative unital ring in which every non-zero element is a unit. That is, 
$$U(F) = F \setminus \{0_{F}\}.$$
\end{definition}

\begin{example}
\begin{itemize}
    \item $\mathbb{Q}, \mathbb{R},$ and $\mathbb{C}$ are fields.
    \item $\mathbb{Z}$ is \textbf{not} a field (the only units are $\pm 1$).
    \item $\mathbb{Z}_n$ is a field if and only if $n$ is prime. If $p$ is prime, we often denote this field as $\mathbb{F}_p$.
    \item $M_{n \times n}(\mathbb{R})$ is \textbf{not} a field (it is not commutative, and not every non-zero matrix is invertible).
\end{itemize}
\end{example}


\section{Ring Homomorphisms}
Just as group homomorphisms preserve the group structure, ring homomorphisms are maps between rings that preserve both addition and multiplication.
\begin{definition}[Ring Homomorphism]
Let $R$ and $S$ be rings. A map $\phi: R \to S$ is a \textbf{ring homomorphism} if for all $a, b \in R$:
\begin{enumerate}
    \item $\phi(a +_R b) = \phi(a) +_S \phi(b)$
    \item $\phi(a \cdot_R b) = \phi(a) \cdot_S \phi(b)$
\end{enumerate}
Furthermore:
\begin{itemize}
    \item If $R$ and $S$ are unital, $\phi$ is \textbf{unital} if $\phi(1_R) = 1_S$.
    \item If $\phi$ is bijective, it is called a \textbf{ring isomorphism}, denoted $R \cong S$.
\end{itemize}
\end{definition}

\begin{example}
\begin{itemize}
    \item Consider $\phi: \mathbb{Z} \to \mathbb{Z}$ given by $\phi(n) := 2n$.
    \begin{enumerate}
        \item $\phi$ is a group homomorphism on $(\mathbb{Z}, +)$.
        \item However, $\phi$ is \textbf{not} a ring homomorphism because for $a, b \neq 0$:
        \[ \phi(a \cdot b) = 2ab \neq 4ab = (2a) \cdot (2b) = \phi(a) \cdot \phi(b) \]
    \end{enumerate} 
    
    \item In $\mathbb{Z}_{10}$, the map $\phi(n) = 2n$ is not a ring homomorphism, but $\phi(n) := 5n$ \textbf{is} a ring homomorphism.
    \begin{itemize}
        \item Addition: $5(a+b) \equiv 5a + 5b \pmod{10}$.
        \item Multiplication: $\phi(ab) = 5ab$ and $\phi(a)\phi(b) = 25ab$. Since $25 \equiv 5 \pmod{10}$, $\phi(ab) = \phi(a)\phi(b)$.
    \end{itemize}

    \item The natural projection $\pi: \mathbb{Z} \to \mathbb{Z}_m$ where $\pi(n) = [n]_m$ is a ring homomorphism.

    \item \textbf{Evaluation Homomorphism:} Let $R$ be a commutative unital ring. Define $\phi: R[x] \to R$ by $\phi(p(x)) := p(1_R)$. This map evaluates a polynomial at the multiplicative identity.
    
    \item The inclusion maps $\mathbb{Z} \hookrightarrow \mathbb{Q}$, $\mathbb{Q} \hookrightarrow \mathbb{R}$, and $\mathbb{R} \hookrightarrow \mathbb{C}$ are ring homomorphisms.
\end{itemize}
\end{example}

\begin{proposition}
Let $\phi: R \to S$ be a ring homomorphism.
\begin{enumerate}
    \item $\phi(0_R) = 0_S$.
    \item $\phi(-a) = -\phi(a)$ for all $a \in R$.
    \item If $\phi$ is unital and $a \in U(R)$, then $\phi(a) \in U(S)$ with $\phi(a)^{-1} = \phi(a^{-1})$.
    \item If $\phi: R \to S$ is an isomorphism, then $\phi^{-1}: S \to R$ is also a ring isomorphism.
\end{enumerate}
\end{proposition}

\begin{proof}
\begin{enumerate}
    \item $\phi(r) = \phi(0_R + r) = \phi(0_R) + \phi(r)$. By the uniqueness of the additive identity in $S$, $\phi(0_R) = 0_S$.
    \item This follows from the properties of group homomorphisms under addition.
    \item $\phi(a^{-1}) \cdot \phi(a) = \phi(a^{-1} a) = \phi(1_R) = 1_S$. Similarly, $\phi(a) \phi(a^{-1}) = 1_S$. Thus, $\phi(a)$ is a unit with inverse $\phi(a^{-1})$.
    \item Since $\phi$ is a bijection, $\phi^{-1}$ exists as a set-theoretic map. Let $\alpha, \beta \in S$. Then there exist unique $a, b \in R$ such that $\phi(a) = \alpha$ and $\phi(b) = \beta$. Then:
    \[ \phi^{-1}(\alpha \cdot \beta) = \phi^{-1}(\phi(a) \cdot \phi(b)) = \phi^{-1}(\phi(ab)) = ab = \phi^{-1}(\alpha) \cdot \phi^{-1}(\beta) \]
    The proof for addition follows the same logic.
\end{enumerate}
\end{proof}

\begin{proposition}
Let $\phi: R \to S$ be a ring homomorphism. Then $\ker \phi := \{ r \in R \mid \phi(r) = 0_S \}$ is a subring of $R$, and $\text{im } \phi := \{ \phi(r) \mid r \in R \}$ is a subring of $S$.
\end{proposition}

\begin{proof}
Let $a, b \in \ker \phi$, so $\phi(a) = \phi(b) = 0_S$. We check the subring criteria for $\ker \phi$:
\begin{itemize}
    \item $\phi(a+b) = \phi(a) + \phi(b) = 0_S + 0_S = 0_S$, so $a+b \in \ker \phi$.
    \item $\phi(-a) = -\phi(a) = -0_S = 0_S$, so $-a \in \ker \phi$.
    \item $\phi(a \cdot b) = \phi(a) \cdot \phi(b) = 0_S \cdot 0_S = 0_S$, so $ab \in \ker \phi$.
\end{itemize}
The proof for $\mathrm{im }\ \phi$ is similar.
\end{proof}


\section{Ideals}
The concept of an ideal is the ring-theoretic analogue of a normal subgroup. Indeed, we are using the same notation $\lhd$ to denote an ideal of a ring:
\begin{definition}[Ideal]
Let $R$ be a ring. A subset $I \lhd R$ is an \textbf{ideal} if:
\begin{enumerate}
    \item $I$ is an additive subgroup of $(R, +)$.
    \item For all $x \in R$ and $i \in I$, we have $xi \in I$ and $ix \in I$.
\end{enumerate}
\end{definition}

\begin{remark}
If $I \triangleleft R$, then $I$ is automatically a subring of $R$. This is because for any $a, b \in I$, the second condition (absorption) implies that $ab \in I$ since $a \in R$ and $b \in I$.

However, not all subrings are ideals. For instance, $\mathbb{Z} \leq \mathbb{Q}$ is a subring but \textbf{not} an ideal, since $2 \in \mathbb{Z}$ and $\frac{1}{3} \in \mathbb{Q}$, but $2 \cdot \frac{1}{3} = \frac{2}{3} \notin \mathbb{Z}$.
\end{remark}

\begin{example}
\begin{itemize}
    
    \item For $R = \mathbb{Z}$, the subset $I = n\mathbb{Z}$ is an ideal. For any $na \in I$ and $m \in \mathbb{Z}$, $(na)m = n(am) \in I$.
    \item Let $R = \mathbb{Z}[x]$ and $I = \{ p(x) \in R \mid p(0) = 0 \}$, the set of polynomials with no constant term. For any $p(x) \in I$ and $q(x) \in R$, the product $p(x)q(x)$ satisfies $p(0)q(0) = 0 \cdot q(0) = 0$, so $p(x)q(x) \in I$.
    \item Let $R = \mathbb{Z}[x]$ and $I = \{ p(x) \in R \mid p(0) \in 2\mathbb{Z} \}$, the set of polynomials where the constant term is an even integer. This is an ideal of $\mathbb{Z}[x]$.
\end{itemize}
\end{example}

\noindent\textbf{Question:}\\
For any ring $R$, how do we construct an ideal $I$?

\begin{definition}[Ideal Generated by $V$]
Let $R$ be a ring and $V \subseteq R$ be a subset. The \textbf{ideal generated by $V$}, denoted $\langle V \rangle$, is the smallest ideal in $R$ containing all elements of $V$.
\end{definition}

\begin{example}
\begin{itemize}
    \item Let $R = \mathbb{Z}$ and $V = \{n\}$. Any ideal $I$ containing $n$ must contain all its additive multiples $nk$ for $k \in \mathbb{Z}$ because $I$ is a subgroup. Thus $n\mathbb{Z} \subseteq I$. Since $n\mathbb{Z}$ is itself an ideal, we have $\langle n \rangle = n\mathbb{Z}$.
    \item In the general case, let $R$ be a unital commutative ring and $V = \{a_1, \dots, a_n\}$. Then:
    \[ \langle a_1, \dots, a_n \rangle = \{ r_1a_1 + \dots + r_na_n \mid r_i \in R \} \]
    To verify this:
    \begin{enumerate}
        \item This set $S$ contains each $a_i$ (e.g., $a_1 = 1_R\cdot a_1 + 0_R \cdot a_2 + \dots$).
        \item $S$ is an ideal: the sum of two such linear combinations is another linear combination, and multiplying by $r \in R$ distributes over the terms.
        \item Any ideal $I$ containing $\{a_1, \dots, a_n\}$ must contain $r_i a_i$ by absorption, and their sum by the subgroup property. Thus $S \subseteq I$.
    \end{enumerate}
    \item Let $R = \mathbb{Z}[x]$. The ideal $\langle 2, x \rangle = \{ 2p(x) + xq(x) \mid p(x), q(x) \in R \}$ is exactly the set of polynomials with an even constant term.
\end{itemize}
\end{example}

\begin{lemma}
Suppose $I_1, \dots, I_k \triangleleft R$. Then:
\begin{enumerate}
    \item $I_1 + \dots + I_k := \{ i_1 + \dots + i_k \mid i_r \in I_r \}$ is an ideal of $R$.
    \item $\bigcap_{i=1}^k I_i$ is an ideal of $R$.
\end{enumerate}
\end{lemma}

\begin{proof}
\begin{enumerate}
    \item Let $x = (i_1 + \dots + i_k)$ and $y = (i_1' + \dots + i_k')$ be elements of the sum. 
    Then $x - y = (i_1 - i_1') + \dots + (i_k - i_k')$. Since each $I_r$ is an additive subgroup, $i_r - i_r' \in I_r$, so $x - y \in I_1 + \dots + I_k$.
    For any $r \in R$, $r(i_1 + \dots + i_k) = ri_1 + \dots + ri_k$. Since $ri_j \in I_j$, the sum is in $I_1 + \dots + I_k$.
    \item Let $x, y \in \bigcap I_i$. Then for all $i$, $x, y \in I_i$. Since each $I_i$ is an ideal, $x-y \in I_i$ and $rx \in I_i$ for all $i$. Therefore, $x-y \in \bigcap I_i$ and $rx \in \bigcap I_i$.
\end{enumerate}
\end{proof}

\begin{example}
In $R = \mathbb{Z}$, let $I_r = \langle a_r \rangle = a_r \mathbb{Z}$ for $a_r \in \mathbb{N}$. Then:
\begin{itemize}
    \item $I_1 + \dots + I_k = \langle \gcd(a_1, \dots, a_k) \rangle = \gcd(a_1, \dots, a_k) \mathbb{Z}$
    \item $I_1 \cap \dots \cap I_k = \langle \text{lcm}(a_1, \dots, a_k) \rangle =\text{lcm}(a_1, \dots, a_k) \mathbb{Z}$
\end{itemize}
Indeed, one can show that if $I \lhd \mathbb{Z}$ is an ideal, then $I = \langle i \rangle$ for some $i \in \mathbb{Z}$ (Homework).
\end{example}

\begin{definition}[Principal Ideal]
Let $R$ be a unital commutative ring. An ideal $I \triangleleft R$ is called \textbf{principal} if $I = \langle a \rangle = \{ar\ |\ r \in R\}$ for some $a \in R$.
\end{definition}
For instance, all ideals in $\mathbb{Z}$ are principal. However, the ideal $\langle 2, x \rangle \lhd \mathbb{Z}[x]$ is not principal - one cannot find any $p(x) \in \mathbb{Z}[x]$ such that $\langle 2,x \rangle = \langle p(x) \rangle$ (Homework).


\begin{proposition}
Let $\phi: R \to S$ be a ring homomorphism. Then $\ker \phi \triangleleft R$.
\end{proposition}

\begin{proof}
From previous results, we know that $\ker \phi$ is a subring of $R$. It remains to show the absorption property: for all $r \in R$ and $i \in \ker \phi$, $ri \in \ker \phi$ and $ir \in \ker \phi$.
Indeed, $\phi(ri) = \phi(r)\phi(i) = \phi(r) \cdot 0_S = 0_S$. Hence, $ri \in \ker \phi$. Similarly, $\phi(ir) = \phi(i)\phi(r) = 0_S \cdot \phi(r) = 0_S$, so $ir \in \ker \phi$. \qedhere
\end{proof}

\section{Quotient Ring}

\begin{definition}[Quotient Ring]
Let $R$ be a ring and $I \triangleleft R$ an ideal. The \textbf{quotient ring} $R/I$ is the set of additive cosets $R/I := \{ r + I \mid r \in R \}$ equipped with the operations:
\begin{align*}
(r_1 + I) + (r_2 + I) &:= (r_1 + r_2) + I \\
(r_1 + I) \cdot (r_2 + I) &:= (r_1 \cdot r_2) + I
\end{align*}
\end{definition}
We need to check the above definition is well-defined, and satisfies the axioms of a ring.
\begin{itemize}
    \item \textbf{Multiplication is well-defined:} 
    Suppose $r_1 + I = r_1' + I$ and $r_2 + I = r_2' + I$. This implies $r_1' = r_1 + i_1$ and $r_2' = r_2 + i_2$ for some $i_1, i_2 \in I$. Then:
    \[ r_1'r_2' = (r_1 + i_1)(r_2 + i_2) = r_1r_2 + r_1i_2 + i_1r_2 + i_1i_2 \]
    Since $I$ is an ideal, $r_1i_2, i_1r_2,$ and $i_1i_2$ are all in $I$. Thus $r_1'r_2' - r_1r_2 \in I$, which means $r_1'r_2' + I = r_1r_2 + I$.
    \item \textbf{Addition is well-defined:} This follows directly from the fact that $(I, +)$ is a normal subgroup of $(R, +)$.
    \item \textbf{Ring Axioms:} Associativity and Distributivity in $R/I$ are inherited from the properties of $R$. For example:
    \[ (r_1+I)((r_2+I) + (r_3+I)) = r_1(r_2+r_3) + I = (r_1r_2 + r_1r_3) + I = (r_1+I)(r_2+I) + (r_1+I)(r_3+I) \]
\end{itemize}

It is easy to see that the following holds for quotient rings:
\begin{itemize}
    \item If $R$ is commutative, then $R/I$ is commutative.
    \item If $R$ is unital, then $R/I$ is unital with identity $1_{R/I} = 1_R + I$.
\end{itemize}
We will leave it to the reader to verify these statements.

\begin{example} \label{eg-zerodiv}
\begin{itemize}
    \item Consider $\mathbb{R}[x]/\langle x^2 - 1 \rangle$. Let $\overline{p(x)} = p(x) + \langle x^2 - 1 \rangle$. 
    In this ring, 
    $$\overline{x-1} \cdot \overline{x+1} = \overline{x^2-1} = \overline{0}.$$ 
    So there exists two nonzero elements such that its product is zero (these elements are called {\bf zerodivisors}, which will be discussed in greater detail in Section \ref{sec-ID} below).
    %Since the product of two non-zero elements is zero, this ring has zerodivisors and is \textbf{not} an integral domain. 
    This occurs because $x^2-1$ is reducible in $\mathbb{R}[x]$.
    
    \item In $\mathbb{R}[x]/\langle x^2 + 1 \rangle$, we have $\overline{x} \cdot \overline{x} = \overline{x^2}$. Since $x^2 \equiv -1 \pmod{x^2+1}$, we have $\overline{x}^2 = \overline{-1}$. This construction behaves like the complex numbers, where $\overline{x}$ plays the role of `square root of $-1$'.
    
    \item Consider $\mathbb{Z}[x]/\langle 2, x \rangle$. Any polynomial $p(x) = a_nx^n + \dots + a_1x + a_0$ is congruent to its constant term $a_0$ modulo the ideal $\langle 2, x \rangle$. Furthermore, $a_0 \equiv 0 \text{ or } 1 \pmod 2$. Thus, 
    $$\mathbb{Z}[x]/\langle 2, x \rangle \cong \mathbb{Z}_2.$$
\end{itemize}
\end{example}

\begin{theorem}[First Isomorphism Theorem of Rings]
Let $\Phi: R \to S$ be a ring homomorphism. Then the map 
$$\psi: R/\ker \Phi \to \mathrm{im}\ \Phi$$ 
defined by $\psi(r + \ker \Phi) := \Phi(r)$ is a ring isomorphism.
\end{theorem}

\begin{proof}
\begin{itemize}
    \item \textbf{Well-defined and Bijective:} From the First Isomorphism Theorem for Groups, we know $\psi$ is a well-defined bijection for the additive structure.
    \item \textbf{Homomorphism:} We check the multiplicative property:
    \[ \psi((r + \ker \Phi)(r' + \ker \Phi)) = \psi(rr' + \ker \Phi) = \Phi(rr') = \Phi(r)\Phi(r') = \psi(r + \ker \Phi)\psi(r' + \ker \Phi). \]
    Since $\psi$ preserves both addition and multiplication and is a bijection, it is a ring isomorphism. \qedhere
\end{itemize}
\end{proof}

\begin{example}
\begin{itemize}
    \item Define $\Phi: \mathbb{R}[x] \to \mathbb{C}$ by $\Phi(p(x)) = p(i)$. This is a surjective homomorphism. The kernel consists of polynomials $p(x)$ such that $p(i) = 0$. By the Factor Theorem, 
    $$(x-i) \mid p(x).$$ 
    Since $p(x) \in \mathbb{R}[x]$ has real coefficients, the complex roots come in pairs, so one also has $(x+i) \mid p(x)$. Thus $(x^2+1) \mid p(x)$, meaning $\ker \Phi = \langle x^2+1 \rangle$. By the First Isomorphism Theorem:
    \[ \mathbb{R}[x]/\langle x^2 + 1 \rangle \cong \mathbb{C} \]
    
    \item Define $\Phi: \mathbb{Z}[x] \to \mathbb{Z}_2$ by $\Phi(p(x)) = p(0) \pmod 2$. The kernel is the set of polynomials with an even constant term, which is the ideal $\langle 2, x \rangle$. By the First Isomorphism Theorem:
    \[ \mathbb{Z}[x]/\langle 2, x \rangle \cong \mathbb{Z}_2 \]
\end{itemize}
\end{example}

\section{Product Ring and Chinese Remainder Theorem}
\begin{definition}[Product Ring]
Given rings $R_1$ and $R_2$, the Cartesian product $R_1 \times R_2$ is a ring under component-wise operations:
\[ (r_1, r_2) + (s_1, s_2) = (r_1 + s_1, r_2 + s_2) \]
\[ (r_1, r_2) \cdot (s_1, s_2) = (r_1 \cdot s_1, r_2 \cdot s_2). \]
We can generalize the above definition to $\prod_{i=1}^n R_i = R_1 \times R_2 \times \dots \times R_n.$
\end{definition}


\begin{definition}
Let $R$ be a unital commutative ring. We say $I_1,I_2\triangleleft R$ \textbf{coprime} if $I_1+I_2=R$.
\end{definition}

\begin{example}
Let $R=\mathbb{Z}$ and $I_1=\langle m\rangle=m\mathbb{Z}$, $I_2=\langle n\rangle=n\mathbb{Z}$. We have seen before that $I_1+I_2:=\{mp+nq\mid p,q\in\mathbb{Z}\}=\langle gcd(m,n)\rangle$. Therefore, 
\begin{center}
$I_1 , I_2$ are coprime $\Leftrightarrow$ $I_1+I_2=\mathbb{Z}$ $\Leftrightarrow$ $\langle \gcd(m,n)\rangle=\mathbb{Z}$ $\Leftrightarrow$ $\gcd(m,n)=1$ are coprime as integers.
\end{center}
\end{example}

The above example generalizes our understanding of "coprime" from $\mathbb{Z}$ to any $R$: 
Two elements $r_1,r_2$ are coprime in $R$ means $\langle r_1\rangle+\langle r_2\rangle=R$.

\begin{theorem}[Chinese Remainder Theorem]
Let $R$ be a commutative unital ring, and let $I_1, \ldots, I_k \triangleleft R$ be ideals such that $I_i$ and $I_j$ are pairwise coprime for $i \neq j$. Then we have a ring isomorphism 
$$\phi: R/(I_1 \cap \cdots \cap I_k) \rightarrow R/I_1 \times \cdots \times R/I_k$$ 
defined by $\phi(r + I_1 \cap \cdots \cap I_k) := (r + I_1, \ldots, r + I_k)$.
\end{theorem}

\begin{proof}
Let $\Phi: R \rightarrow R/I_1 \times \cdots \times R/I_k$ be the ring homomorphism defined by 
$$\Phi(r) := (r + I_1, \ldots, r + I_k).$$
By the First Isomorphism Theorem, we need to show:
\begin{itemize}
\item[(i)] $\ker\Phi = I_1 \cap \cdots \cap I_k$, and 
\item[(ii)] $\operatorname{im}\Phi = R/I_1 \times \cdots \times R/I_k$.
\end{itemize}

\textbf{Part (i):} We have
\begin{align*}
r \in \ker\Phi 
&\Leftrightarrow \Phi(r) = (0 + I_1, \ldots, 0 + I_k) \\
&\Leftrightarrow r + I_\ell = 0 + I_\ell \text{ for all } \ell = 1, \ldots, k \\
&\Leftrightarrow r \in I_\ell \text{ for all } \ell = 1, \ldots, k \\
&\Leftrightarrow r \in I_1 \cap \cdots \cap I_k.
\end{align*}

\textbf{Part (ii):} Fix $j \in \{1, \ldots, k\}$. Since $R$ is unital and $I_j$ and $I_i$ are coprime for all $i \neq j$, we have $I_i + I_j = R$. Thus there exist $z_i \in I_j$ and $w_i \in I_i$ such that 
$$z_i + w_i = 1 \quad \text{for all } i \neq j.$$

Consider the identity $1 = (1 - \prod_{i \neq j} w_i) + \prod_{i \neq j} w_i$. Define
$$x_j := 1 - \prod_{i \neq j} w_i = 1 - \prod_{i \neq j} (1 - z_i).$$
Expanding the product, we see that $x_j$ is a sum of products of the $z_i$'s with no constant term, hence $x_j \in I_j$. Also, define
$$y_j := \prod_{i \neq j} w_i \in \bigcap_{i \neq j} I_i,$$
since each $w_i \in I_i$ and ideals are closed under multiplication.

Therefore, for each fixed $j$, we have 
$$1 = x_j + y_j, \quad \text{where } x_j \in I_j \text{ and } y_j \in \bigcap_{i \neq j} I_i.$$

We now verify that $\Phi$ is surjective. Let $(u_1 + I_1, \ldots, u_k + I_k) \in R/I_1 \times \cdots \times R/I_k$ be an arbitrary element. 

\textbf{Claim:} $\Phi(u_1 y_1 + \cdots + u_k y_k) = (u_1 + I_1, \ldots, u_k + I_k)$.

\textbf{Proof of Claim:} We have 
$$\Phi(u_1 y_1 + \cdots + u_k y_k) = (u_1 y_1 + \cdots + u_k y_k + I_1, \ldots, u_1 y_1 + \cdots + u_k y_k + I_k).$$
For the $\ell$-th component, note that $y_j \in I_\ell$ for all $j \neq \ell$ (by definition of $y_j$), so 
\begin{align*}
u_1 y_1 + \cdots + u_k y_k + I_\ell 
&= u_\ell y_\ell + I_\ell \\
&= u_\ell(1 - x_\ell) + I_\ell \\
&= u_\ell - u_\ell x_\ell + I_\ell \\
&= u_\ell + I_\ell,
\end{align*}
where the last equality holds because $x_\ell \in I_\ell$.

Thus $\Phi$ is surjective, completing the proof.
\end{proof}

\begin{corollary}
Let $p_1,\cdots,p_n$ be distinct prime numbers. Then $$\mathbb{Z}/\langle p_1^{a_1}\cdots p_n^{a_n}\rangle\cong\mathbb{Z}/\langle p_1^{a_1}\rangle\times\cdots\times\mathbb{Z}/\langle p_n^{a_n}\rangle.$$
\end{corollary}

Therefore, for each $b_i\in \mathbb{Z}/\langle p_i^{a_i}\rangle\cong \mathbb{Z}_{p_i^{a_i}}$, there exists $x\in \mathbb{Z}$ such that 
$\Phi(x)=(b_1+\langle p_1^{a_1}\rangle,\cdots,b_n+\langle p_n^{a_n}\rangle)$, i.e. 
$$x\equiv b_i \pmod{\mathbb{Z}_{p_i^{a_i}}}$$
for all $i$.



\section{Integral Domains} \label{sec-ID}

In many rings, the product of two non-zero elements can be zero. Integral domains are a class of rings where this ``pathological'' behavior does not occur, much like in the integers.

\begin{definition}[Zerodivisor]
Let $R$ be a ring. A non-zero element $r \in R$ is called a \textbf{zerodivisor} if there exists a non-zero element $s \in R$ such that $r \cdot s = 0$ or $s \cdot r = 0$.
\end{definition}

\begin{example}
\begin{itemize}
\item In the ring $R = \mathbb{Z}_6$, the element $[2]$ is a zerodivisor because $[2] \cdot [3] = [6] = [0]$, and both $[2], [3] \neq [0]$. 

\item Let $R$ be unital commutative ring, and $R[x]$ be its polynomial ring, and let $p(x) = a(x)b(x) \in R[x]$ be a reducible polynomial. Then
$$\overline{a(x)} := a(x) + \langle p(x) \rangle, \quad \overline{b(x)} := b(x) + \langle p(x) \rangle$$
are zerodivisors of the quotient ring $R[x]/\langle p(x) \rangle$ (c.f. Example \ref{eg-zerodiv}).

\item Finally, if $R_1, R_2$ are not zero rings, then the product ring $R_1 \times R_2$ has zerodivisors. Namely, $$(r_1,0_{R_2}) \cdot (0_{R_1},r_2) = (0_{R_1},0_{R_2}) = 0_{R_1 \times R_2}$$
for $r_1 \neq 0$, $r_2 \neq 0$. So they are both zerodivisors.
\end{itemize}
\end{example}


As one noticed in the construction of ideal $\langle a_1, \dots, a_r \rangle \lhd R$ generated by $a_1, \dots, a_r \in R$ and the Chinese Remainder Theorem, unital commutative rings have a much nicer structure than general rings. 
Therefore, we separate the case when $R$ is unital commutative in the definition below:
\begin{definition}[Integral Domain]
A ring $R$ is a \textbf{domain} if it has no zerodivisors. A {\it unital, commutative} ring with no zerodivisors is called an \textbf{integral domain} (ID).
\end{definition}


\begin{example}
\begin{itemize}
    \item $\mathbb{Z}, \mathbb{Q}, \mathbb{R}, \mathbb{C},$ and the Gaussian integers $\mathbb{Z}[i]$ are all integral domains.
    \item $\mathbb{Z}_n$ is an integral domain if and only if $n$ is prime. If $n$ is composite, say $n = ab$ with $1 < a, b < n$, then $[a][b] = [0]$ in $\mathbb{Z}_n$, so it has zerodivisors.
    \item If $R$ is an integral domain, then the polynomial ring $R[x]$ is also an integral domain.
    %\item The matrix ring $M_{n \times n}(\mathbb{R})$ for $n \ge 2$ is \textbf{not} an integral domain, as it is non-commutative and contains zerodivisors (e.g., matrices with only one non-zero entry in different positions).
\end{itemize}
\end{example}

\begin{proposition}[Cancellation Property]
Let $R$ be a (unital, commutative) ring. Then $R$ is a(n integral) domain if and only if for all $a, b, c \in R$ with $c \neq 0$:
\[ ca = cb \implies a = b \]
\end{proposition}

\begin{proof}
$(\Rightarrow)$ Suppose $R$ is an integral domain and $ca = cb$ with $c \neq 0$. Then $ca - cb = 0$, which implies $c(a - b) = 0$. Since $R$ has no zerodivisors and $c \neq 0$, we must have $a - b = 0$, hence $a = b$.

$(\Leftarrow)$ Suppose the cancellation property holds. Let $c, x \in R$ such that $cx = 0$ and $c \neq 0$. We can write this as $cx = c \cdot 0$. By cancellation, $x = 0$. Thus, $R$ has no zerodivisors.
\end{proof}

\begin{remark}
The cancellation property is the primary reason integral domains are `well-behaved'. It allows us to solve algebraic equations similarly to how we do in the integers or fields.
\end{remark}

\noindent\textbf{Hierarchy of Rings:} \\
One of the main goals of ring theory is to understand the following inclusions:
\begin{align*} \text{Fields} &\subsetneq \text{Euclidean Domains (ED)}\\ &\subsetneq \text{Principal Ideal Domains (PID)} \\ &\subsetneq \text{Unique Factorization Domain (UFD)} \\ &\subsetneq \text{Integral Domains (ID)} \end{align*}

Indeed, one can easily see the coarsest inclusion holds:
\begin{proposition}
If $R$ is a field, then $R$ is an integral domain.
\end{proposition}

\begin{proof}
Since $R$ is a field, it is a commutative ring with unit $1_R$. Suppose on contrary that there exist $a,b \in R$ such that $ab = 0_R$ with $a, b \neq 0_R$. Since $R$ is a field and $a \neq 0_R$, the element $a$ has a multiplicative inverse $a^{-1}$. Multiplying both sides of the equation $ab = 0$ by $a^{-1}$ gives
\[
a^{-1}(ab) = a^{-1}0_R.
\]
By associativity,
\[
(a^{-1}a)b = 0_R.
\]
Since $a^{-1}a = 1_R$, we obtain
\[
1_R \cdot b = 0_R,
\]
which implies $b = 0_R$, contradicting the assumption that $ab = 0_R$ with $a, b \neq 0_R$. Therefore, $R$ has no zero divisors, and $R$ is an integral domain.
\end{proof}

In the case when $R$ is a finite ring, the converse holds:
\begin{lemma}
Let $R$ be a finite domain. Then $R$ is unital.
\end{lemma}

\begin{proof}
Let $a \in R$ be a non-zero element. Consider the set of powers $\{a, a^2, a^3, \dots\}$. Since $R$ is finite, there must be repetitions: $a^m = a^n$ for some $m > n > 0$. By the cancellation property, $a^{m-n} a^n = 1 \cdot a^n$, so $a^{m-n}$ acts as the identity for powers of $a$. 

Specifically, for any $x \in R$, consider $x a^n$. We have $x a^m = x a^n a^{m-n}$. Since $a^m = a^n$, this implies $x a^n = (x a^{m-n}) a^n$. By cancellation of $a^n$, we get $x = x a^{m-n}$. Thus $a^{m-n}$ is the multiplicative identity $1_R$.
\end{proof}

\begin{proposition}
Every finite commutative domain $R$ is a field.
\end{proposition}

\begin{proof}
Let $R$ be a finite commutative domain, and let $R \setminus \{0_R\} = \{r_1, r_2, \dots, r_m\}$. By the Lemma above, we have seen that $R$ is unital.
To further show $R$ is a field, we must show every non-zero $a \in R$ has a multiplicative inverse. 

Consider the set $S = \{ar_1, ar_2, \dots, ar_m\}$. If $ar_i = ar_j$, then by the cancellation property $r_i = r_j$, so $i = j$. Thus, all $m$ elements in $S$ are distinct and non-zero. Since $S \subseteq R \setminus \{0_R\}$ and $|S| = |R \setminus \{0_R\}|$, it must be that $S = R \setminus \{0_R\}$. 

Since the identity $1_R$ is in $R \setminus \{0\}$, there must exist some $r_k$ such that $ar_k = 1_R$. Thus $a$ is a unit, and $R$ is a field.
\end{proof}


For the rest of this chapter, we will further focus on the case when {\bf $R$ is an (infinite) integral domain}, and study the intermediate rings in the above inclusions.









\section{Prime and Maximal Ideals}

As stated in the previous section, one would like to study various arithmetic properties for integral domain $R$. 

To facilitate for generalization, we would first define these properties in terms of {\bf ideals}, and then derive the equivalent conditions in terms of {\bf elements}.
\begin{definition}[Divides and Associates]
Let $a,b \in R$, where $R$ is an integral domain.  
We say that $a$ \textbf{divides} $b$ (written $a \mid b$) if
\[
\langle b \rangle \subseteq \langle a \rangle.
\]
(or equivalently, \quad \( b \in \langle a \rangle\) \quad $\Leftrightarrow$ \quad  $b = qa$ for some $q \in R$.). 

We say that $a$ and $b$ are \textbf{associates}, written $a \sim b$, if \(a \mid b \ \text{and} \ b \mid a\). Or equivalently,
\[
\langle a \rangle = \langle b \rangle.
\]
\end{definition}
To understand these definitions purely in terms of elements, one has:
\begin{lemma}
Let $R$ be an integral domain and let $a,b \in R$. Then
\begin{center}
$a \sim b$
if and only if there exists $u \in U(R)$ such that $a = ub$.
\end{center}
\end{lemma}

\begin{proof}
($\Leftarrow$) Suppose $a = ub$ for some unit $u \in U(R)$.  
Then $a \in \langle b \rangle$, so $\langle a \rangle \subseteq \langle b \rangle$. Since $u$ is a unit, $u^{-1}$ exists, and
\[
b = u^{-1}a,
\]
so $b \in \langle a \rangle$, hence $\langle b \rangle \subseteq \langle a \rangle$ and therefore $\langle a \rangle = \langle b \rangle$. In other words, $a \sim b$.

\medskip

($\Rightarrow$) Suppose $a \sim b$, i.e. $\langle a \rangle = \langle b \rangle$. Then $a \in \langle b \rangle$, so there exists $p \in R$ such that
\(
a = pb.
\) Similarly, since $b \in \langle a \rangle$, there exists $q \in R$ such that
\(
b = qa.
\)

Substituting gives
\[
a = p(qa) = (pq)a.
\]

Since $R$ is an integral domain and $a \neq 0$, we may cancel $a$ to obtain
\[
1 = pq.
\]

Thus $p$ is a unit (with inverse $q$), and
\[
a = pb
\]
with $p \in U(R)$. Hence $a = ub$ for some unit $u$.
\end{proof}




Now we define what `prime' really means for $R$.
Rather than studying whether an element $r\in R$ is prime directly, we define what it means for an ideal $I\triangleleft R$ to be prime, and then define what it means for $p \in R$ to be prime. As an example, recall that we defined


%To better understand these new notions in terms of ideals, it is recommended to go to the basic case when $R=\mathbb{Z}$. For instance, to understand the statement
%\begin{center}
%$I_1, I_2, \dots \triangleleft R$\quad satisfy property $(*)$
%\end{center}
%(here $(*)$ can be prime, irreducible, coprime, $\dots$), try to understand the property in the special case:
%\begin{center}
%$I_1 = \langle n_1 \rangle ,\ I_2 = \langle n_2 \rangle,\ \dots \triangleleft \mathbb{Z}$\quad satisfy property $(*)$
%\end{center}
%and further convert the definition of the property to relations between the elements $n_1, n_2, \dots$ in $\mathbb{Z}$. 


\begin{center}
    $I_1$, $I_2$ is coprime if $I_1 + I_2 = R$ \quad \quad (coprime ideals)
\end{center}
And we define:
\begin{definition}[Coprime elements]
    Let $R$ be a unital commutative ring. Two elements $a$, $b$ are {\bf coprime} if the principal ideals $\langle a\rangle$, $\langle b \rangle \lhd R$ are coprime, i,e,
    $$\langle a \rangle + \langle b \rangle = R.$$
\end{definition}
To better comprehend the above definition, the above condition is equivalent to saying:
\begin{center}
    $a, b \in R$ are coprime\ \ $\Longleftrightarrow$\ \ there exists $p, q \in R$ such that $ap + bq = 1$ (Exercise).
\end{center}
{\bf Upshot: We defined what it means for $a, b \in R$ to be coprime without defining `common factors'!}

Note that in the prototype case when $R=\mathbb{Z}$, the above statement and Bezout's theorem imply that 
\begin{center}
$a, b \in \mathbb{Z}$ are coprime\ \ $\Longleftrightarrow$\ \ $\gcd(a,b)=1$.
\end{center}
which gives us back our usual definition of what coprime means in $\mathbb{Z}$. 

\begin{definition}[Prime and Maximal Ideals]
Let $R$ be an integral domain. We say a {\it proper} ideal (i.e. $I\neq R$) $I\triangleleft R$ is 
\begin{enumerate}
    \item \textbf{prime} if for all $a,b \in R$ such that $ab \in R$, then we must have $a\in I$ or $b\in I$;
    \item \textbf{maximal} if for all ideals $J\triangleleft R$ s.t. $I\subset J\subset R$, then $J=I$ or $J=R$.
\end{enumerate}
\end{definition}

\begin{definition}[Prime Element]
Let $R$ be a unital commutative ring. An element $p \in R$ is called 
\textbf{prime} if the ideal $\langle p \rangle$ is a prime ideal.

Equivalently, $p$ is prime if and only if
\[
p \mid ab \;\Longrightarrow\; p \mid a \text{ or } p \mid b
\]
for all $a,b \in R$.
\end{definition}




\begin{example}
\begin{itemize}
    \item Back into our prototype $R=\mathbb{Z}$. 
    It is obviously true that if $p \in \mathbb{Z}$ is a prime number, then $p\mid ab\ \Leftrightarrow \ p\mid a\ or\; p\mid b$ and vice versa.
    
    (\textbf{Exercise:} $\langle n\rangle$ is a maximal ideal $\Leftrightarrow$ $n$ is prime $\Leftrightarrow$ $\langle n\rangle$ is prime ideal)
    %(\textbf{Nonexample:} $\langle 6\rangle$ is \textbf{NOT} maximal since $\langle 6\rangle \subsetneq \langle 2\rangle\subsetneq \mathbb{Z}$)
%    \item Let $R=\mathbb{Z}_{12}$. All the ideals of $R$ are:
%    \begin{itemize}
%        \item $I_0=\{0\}$ (neither maximal nor prime by definition);
%        \item $\langle 2\rangle =\{0,2,4,6,8,10\}$ (maximal and prime);
%        \item $\langle 3 \rangle=\{0,3,6,9\}$ (maximal and prime);
%        \item $\langle 4 \rangle=\{0,4,8\}$ (neither maximal nor prime);
%        \item $\langle 6 \rangle=\{0,6\}$ (neither maximal nor prime);
%        \item $R$ (neither maximal nor prime by definition) 
%    \end{itemize}
    
    \item Let $R=\mathbb{Z}[x]$, and $I=\langle x\rangle$ be all polynomials with $0$ constant term is a prime ideal -- 
    take 
    $$p(x)=a_0+a_1x+\cdots+a_mx^m,\quad  q(x)=b_0+b_1x+\cdots+b_nx^n$$ 
    Then 
    $$p(x)q(x)\in I\ \Leftrightarrow\ a_0b_0=0\ \Leftrightarrow a_0=0\; or \; b_0=0\Leftrightarrow\ p(x)\in I\;or\; q(x)\in I$$
    However, $I$ is \textbf{NOT} maximal, since
    $$I\subsetneq \langle 2,x\rangle\subsetneq R.$$
\end{itemize}
\end{example}

\begin{proposition}
Let $R$ be an integral domain, and let $I \triangleleft R$. Then:
\begin{enumerate}
    \item $I$ is prime $\Longleftrightarrow R/I$ is an integral domain.
    \item $I$ is maximal $\Longleftrightarrow R/I$ is a field.
\end{enumerate}
\end{proposition}

\begin{example}
Consider the ideal $\langle x \rangle \subset \mathbb{Z}[x]$. Then
\[
\mathbb{Z}[x]/\langle x \rangle \cong \mathbb{Z}.
\]
Since $\mathbb{Z}$ is an integral domain, it follows that $\langle x \rangle$ is prime. However, $\mathbb{Z}$ is \emph{not} a field, hence $\langle x \rangle$ is not maximal.
\end{example}

\begin{proof}
\begin{enumerate}
    \item 
    Let $a+I, b+I \in R/I$. Then
    \(
    (a+I)(b+I) = 0+I
    \Longleftrightarrow
    ab \in I.
    \)

    $(\Rightarrow)$ Suppose $I$ is prime. Suppose $a+I, b +I$ be such that $(a+I)(b+I) = 0 +I$. By the first sentence of the proof, this implies
    \[
    ab \in I \Longrightarrow a \in I \text{ or } b \in I.
    \]
    Hence
    \[
    a+I=0 \text{ or } b+I=0.
    \]
    Therefore $R/I$ has no zero divisors and is an integral domain.

    $(\Leftarrow)$ Conversely, suppose $R/I$ is an integral domain. Consider $ab \in I$. Then the first sentence of the proof says
    \[
    (a+I)(b+I)=0 +I.
    \]
    Since $R/I$ has no zero divisors, this implies
    \[
    a+I=0+I \text{ or } b+I=0+I,
    \]
    hence $a \in I$ or $b \in I$. Thus $I$ is prime.

    \item 
    Suppose $I$ is maximal and let $J$ be an ideal such that
    \[
    I \subseteq J \subseteq R.
    \]
    Then $J = I$ or $J = R$.

    By the Correspondence Theorem (Homework), ideals $J$ with $I \subseteq J \subseteq R$ correspond bijectively to ideals of $R/I$ via
    \[
    J \longmapsto J/I.
    \]
    Hence the only ideals of $R/I$ are $\{0\}$ and $R/I$. 

    Recall (from Homework) that a ring $\mathbb{F}$ is a field if and only if its only ideals are $\{0\}$ and $\mathbb{F}$. Therefore $R/I$ is a field. Conversely, if $R/I$ is a field, then its only ideals are $\{0\}$ and $R/I$. By correspondence theorem again, the only ideals of $R$ containing $I$ are $I$ and $R$. Hence $I$ is maximal.
\end{enumerate}
\end{proof}

\begin{corollary}
Let $R$ be an integral domain. If $I \triangleleft R$ is maximal, then $I$ is prime.
\end{corollary}
\begin{proof}
If $I$ is maximal, then $R/I$ is a field. Since every field is an integral domain, $R/I$ is an integral domain.  
Hence $I$ is prime.
\end{proof}
Indeed, one can proof the above corollary directly without using quotient rings. We leave it as an exercise.


\section{Principal Ideal Domains}

Recall that for any integral domain (indeed for unital commutative rings as well),
\[
I \text{ maximal } \Longrightarrow I \text{ prime}.
\]
However, the converse is \emph{false} in general -- for example, $\langle x\rangle \subset \mathbb{Z}[x]$ is prime but not maximal.

\medskip
A natural goal is to find classes of integral domains $R$ where prime ideals are ``as close as possible'' to maximal ideals. One particularly important class is when \emph{every} ideal is generated by one element.

\begin{definition}[Principal Ideal Domain]
Let $R$ be an integral domain. We say that $R$ is a \textbf{principal ideal domain (PID)} if every ideal
$I \triangleleft R$ is principal, i.e. there exists $a\in R$ such that
\[
I = \langle a \rangle = \{ar \ |\ r\in R\}.
\]
\end{definition}

\begin{example}\hfill
\begin{enumerate}
    \item {\bf All fields $\mathbb{F}$ are PIDs} -- By Homework, the only ideals in $\mathbb{F}$ are $\{0_{\mathbb{F}}\} = \langle 0_{\mathbb{F}} \rangle$ and $\mathbb{F} = \langle 1_{\mathbb{F}} \rangle$.
    \item \textbf{$\mathbb{Z}$ is a PID} -- Let $I \triangleleft \mathbb{Z}$ be a nonzero ideal. Choose $d>0$ to be the smallest positive integer in $I$. \\
    \underline{Claim:} $I=\langle d\rangle$. \\
    Clearly $\langle d\rangle \subseteq I$ since $d\in I$. For the other inclusion, let $n\in I$.
    By the division algorithm in $\mathbb{Z}$, write $n=qd+r$ with $0\le r<d$.
    Then $r=n-qd\in I$. By minimality of $d$, we must have $r=0$, hence $n\in \langle d\rangle$.
    Therefore $I\subseteq \langle d\rangle$, so $I=\langle d\rangle$.

    \item \textbf{If $\mathbb{F}$ is a field, then $\mathbb{F}[x]$ is a PID} -- Let $I\triangleleft \mathbb{F}[x]$ be a nonzero ideal. Choose a nonzero polynomial $f(x)\in I$ of \emph{minimal degree}.
    (Multiplying by a unit in $\mathbb{F}$ if necessary, we may assume $f$ is monic.) \\
    \underline{Claim:} $I=\langle f(x)\rangle$. \\
    Clearly $\langle f\rangle\subseteq I$. For the other inclusion, let $g(x)\in I$.
    By the division algorithm in $\mathbb{F}[x]$, write
    \[
    g(x)=q(x)f(x)+r(x),\qquad \deg r < \deg f.
    \]
    Then $r(x)=g(x)-q(x)f(x)\in I$. By minimality of $\deg f$, we must have $r(x)=0$.
    Hence $g(x)\in \langle f\rangle$, so $I\subseteq \langle f\rangle$ and therefore $I=\langle f\rangle$.

    \item \textbf{$\mathbb{Z}[x]$ is \underline{not} a PID} --
    In particular, the ideal $J=\langle 2,x\rangle \triangleleft \mathbb{Z}[x]$ is not principal (Homework). %If $J=\langle f(x)\rangle$ were principal, then $f(x)\mid 2$, hence $f(x)$ must be a constant $\pm 1,\pm 2$. But $x\in J$ forces $f(x)\mid x$, which is impossible for $f(x)=\pm 2$, and if $f(x)=\pm 1$ then $J=\mathbb{Z}[x]$, contradicting that $J$ is proper. Thus $J$ is not principal.
\end{enumerate}
\end{example}

\begin{remark}
The proofs that $\mathbb{Z}$ and $\mathbb{F}[x]$ are PIDs are the same pattern:
choose an element of \emph{minimal size} (smallest positive integer / smallest degree polynomial) and use the division algorithm to show it generates the entire ideal.

Indeed, these rings belong to a more restrictive class of Integral Domains, called {\bf Euclidean domains (EDs)}. We will study EDs in full detail later, and will show that all EDs are PIDs.
\end{remark}

In a PID, prime and maximal ideals are very close.  Note that since a PID is an integral domain, the zero ideal $\langle 0\rangle$
is always prime (because $R/\langle 0\rangle \cong R$ is an integral domain).  Therefore ``prime $\Leftrightarrow$ maximal'' cannot hold for
\emph{all} ideals unless $R$ is a field.  The correct statement is for \emph{nonzero} prime ideals.

\begin{proposition}
Let $R$ be a PID and let $I\triangleleft R$ be a nonzero proper ideal. Then
\[
I \text{ is prime } \Longleftrightarrow I \text{ is maximal}.
\]
Equivalently, for $0\neq p\in R$ nonunit,
\[
\langle p\rangle \text{ is prime } \Longleftrightarrow \langle p\rangle \text{ is maximal}.
\]
\end{proposition}

\begin{proof}
($\Leftarrow$) Maximal $\Rightarrow$ prime holds in every unital commutative ring.

\medskip
($\Rightarrow$) Assume $I$ is a nonzero prime ideal. Since $R$ is a PID, $I=\langle p\rangle$ for some $0\neq p\in R$.
To show $\langle p\rangle$ is maximal, let $J$ be any ideal with
\[
\langle p\rangle \subseteq J \subseteq R.
\]
Again using that $R$ is a PID, write $J=\langle a\rangle$ for some $a\in R$.
The inclusion $\langle p\rangle \subseteq \langle a\rangle$ means $p\in \langle a\rangle$, i.e. $p=ab$ for some $b\in R$.

Now we use the fact that $\langle p \rangle$ is prime ideal, or equivalently $p \in R$ is prime element. One has
$$p = ab\ \Rightarrow\ p \ | \ ab \ \Rightarrow\  p \ |\ a\ \ \mathrm{or}\ \ p \ |\ b.$$
Suppose first of all $p\ |\ a$. Then $a = pq$ for some $q \in R$. Hence
$$p = ab = pqb \ \Rightarrow\ 1 = qb,$$
i.e. $b \in U(R)$. Consequently
$$p = ab,\ b \in U(R) \ \Rightarrow\ p \sim a \ \Rightarrow \ \langle p \rangle = \langle a \rangle = J$$

On the other hand, suppose $p\ |\ b$. Then one has $a \in U(R)$ similarly, and hence $J = \langle a \rangle = R$. Either way, there is no ideal $J$ lying strictly between $I = \langle p \rangle$ and $R$, so $I$ is maximal.
\end{proof}


\section{Irreducible Elements}

Recall that in a PID, we have shown that nonzero prime ideals are maximal. A further goal is to understand how elements of an integral domain factor. The key notion is that of an \emph{irreducible element}, which serves as the ``building block'' for factorization.

\begin{definition}[Irreducible Element]
Let $R$ be a unital commutative ring. An element $a \in R$ is \textbf{irreducible} if $\langle a \rangle$ is maximal among all proper principal ideals of $R$, i.e., 
\begin{center}
whenever $\langle a \rangle \subseteq \langle b \rangle \subseteq R$ for some $b \in R$, we have $\langle a \rangle = \langle b \rangle$ or $\langle b \rangle = R$.
\end{center}
\end{definition}

As before, we give a characterization of irreducibility in terms of elements of $R$.
\begin{lemma}
Let $R$ be an integral domain, and let $a \in R$ be a nonzero nonunit. Then
\[
a \text{ is irreducible} \iff \text{whenever } a = xy,\ \text{we have } a \sim x \text{ or } a \sim y.
\]
\end{lemma}

\noindent (Recall that $a \sim x$ means $a = ux$ for some unit $u \in U(R)$. In other words, $a$ is irreducible iff whenever $a = xy$, either $x$ is a unit or $y$ is a unit.)

\begin{proof}
$(\Rightarrow)$: Suppose $a$ is irreducible and $a = xy$. Then $a \in \langle x \rangle$, so $\langle a \rangle \subseteq \langle x \rangle$. By irreducibility, either $\langle a \rangle = \langle x \rangle$ or $\langle x \rangle = R$.
\begin{itemize}
    \item If $\langle a \rangle = \langle x \rangle$, then $a \sim x$.
    \item If $\langle x \rangle = R$, then $x$ is a unit. Since $a = xy$ and $x$ is a unit, we have $a \sim y$.
\end{itemize}

$(\Leftarrow)$: Suppose that whenever $a = xy$ we have $a \sim x$ or $a \sim y$. Let $\langle a \rangle \subseteq \langle b \rangle$ for some $b \in R$. Then $a = b b'$ for some $b' \in R$, so by hypothesis, $a \sim b$ or $a \sim b'$.
\begin{itemize}
    \item If $a \sim b$, then $\langle a \rangle = \langle b \rangle$.
    \item If $a \sim b'$, write $a = ub'$ for some unit $u$. Then from $a = bb'$ and $b' = u^{-1}a$, we get $a = b \cdot u^{-1} a$, so $b u^{-1} = 1$ by cancellation in $R$, meaning $b$ is a unit and $\langle b \rangle = R$. \qedhere
\end{itemize}
\end{proof}

\begin{remark}
As an immediate consequence of the lemma, $a \in R$ is irreducible if and only if there do \emph{not} exist nonunits $b, c \in R$ such that $a = bc$. In other words, irreducible elements are precisely the ``atoms'' of $R$: they cannot be factored into strictly smaller pieces. This makes irreducible elements the natural building blocks for factorization in $R$.
\end{remark}

The following proposition shows that prime elements are always irreducible.

\begin{proposition}
Let $R$ be an integral domain. If $0 \neq a \in R$ is prime, then $a$ is irreducible.
\end{proposition}

\begin{proof}
Suppose $a = xy$. Then $a \mid xy$, and since $a$ is prime, $a \mid x$ or $a \mid y$. Without loss of generality, assume $a \mid x$, so $x = a a'$ for some $a' \in R$. Then $a = xy = a a' y$, and cancelling $a$ (which is valid since $R$ is an integral domain and $a \neq 0$) gives $a'y = 1$, so $y$ is a unit. Hence $a \sim x$.
\end{proof}

\begin{remark}
The converse is \emph{false} in general. For example, $3 \in \mathbb{Z}[\sqrt{-5}]$ is irreducible but not prime. However, we will see below that in a PID — and more generally in a UFD — the two notions coincide for nonzero nonunits.
\end{remark}

In a PID, the interplay between maximal ideals, prime ideals, and irreducible elements gives the following clean picture.

\begin{proposition}
Let $R$ be a PID and let $0 \neq a \in R$ be a nonunit. Then
\[
\langle a \rangle \text{ is maximal} \iff \langle a \rangle \text{ is prime} \iff a \text{ is irreducible} \iff a \text{ is prime.}
\]
\end{proposition}

\begin{proof}
The equivalence $\langle a \rangle$ maximal $\Leftrightarrow$ $\langle a \rangle$ prime was established in the previous section. The implication prime $\Rightarrow$ irreducible holds in any integral domain by the proposition above.

It remains to show that irreducible $\Rightarrow$ prime in a PID. Let $a \in R$ be irreducible. By definition, $\langle a \rangle$ is maximal among proper principal ideals. Since every ideal in a PID is principal, $\langle a \rangle$ is in fact a maximal ideal, hence a prime ideal, so $a$ is prime.
\end{proof}

To conclude, the four conditions — maximal ideal, prime ideal, prime element, irreducible element — are all equivalent (for nonzero nonunits) for PIDs. This is a very special property: for instance, in $\mathbb{Z}[x]$ (which is \emph{not} a PID), the element $x$ generates a prime ideal $\langle x \rangle$ that is \emph{not} maximal; and in $\mathbb{Z}[\sqrt{-5}]$ (which is not even a UFD), the element $3$ is irreducible but \emph{not} prime.

\section{Unique Factorization Domain}

We now make precise the notion that every element can be written as a finite product of irreducibles.

\begin{definition}[Factorization Domain]
Let $R$ be an integral domain. We say $R$ is a \textbf{factorization domain} if for every nonzero nonunit $x \in R$, there exist irreducible elements $x_1, \dots, x_r \in R$ such that
\[
x = x_1 x_2 \cdots x_r.
\]
\end{definition}

To prove that PIDs are factorization domains, we first establish a chain condition.

\begin{definition}[ACCP]
Let $R$ be an integral domain. We say $R$ satisfies the \textbf{ascending chain condition on principal ideals} (ACCP) if every ascending chain of principal ideals
\[
\langle a_1 \rangle \subseteq \langle a_2 \rangle \subseteq \langle a_3 \rangle \subseteq \cdots
\]
stabilizes, i.e., there exists $N \in \mathbb{N}$ such that $\langle a_n \rangle = \langle a_N \rangle$ for all $n \geq N$.
\end{definition}

\begin{lemma}
Every PID satisfies the ACCP.
\end{lemma}

\begin{proof}
Let $\langle a_1 \rangle \subseteq \langle a_2 \rangle \subseteq \cdots$ be an ascending chain of principal ideals in a PID $R$. Set $I := \bigcup_{n=1}^{\infty} \langle a_n \rangle$. One verifies that $I$ is an ideal of $R$ (see Homework), so since $R$ is a PID, $I = \langle r \rangle$ for some $r \in R$. Since $r \in I$, we have $r \in \langle a_M \rangle$ for some $M \in \mathbb{N}$. Then for all $n \geq M$,
\[
I = \langle r \rangle \subseteq \langle a_M \rangle \subseteq \langle a_n \rangle \subseteq I,
\]
so $\langle a_n \rangle = I$ for all $n \geq M$, and the chain stabilizes.
\end{proof}

\begin{lemma}
If $R$ satisfies the ACCP, then $R$ is a factorization domain.
\end{lemma}

\begin{proof}
Let 
$$F:=\{x \in R \mid x \sim x_1 \cdots x_r, \ x_i \text{ irreducibles}, r \in \mathbb{N} \cup\{0\}\}
$$ 
We want to show that $R = F$. Note that
\begin{itemize}
\item[(a)] $1 \in F$. More generally, all units are in $F$ ($r = 0$).
\item[(b)] All irreducible elements $x \in R$ are in $F$ ($r=1$).
\item[(c)] $F$ is closed under multiplication (i.e. if $a, b \in F$, then $ab \in F$).
\end{itemize}
Suppose on contrary there exists $x_0 \in R \backslash F$. Then by (b) $x_0 =y_0z_0$ is reducible, where $y_0$, $z_0$ are not units. 
Therefore,
$$x_0 \nsim y_0 \ \mathrm{and} \ x_0 \nsim z_0 \iff \langle x_0 \rangle \neq \langle y_0 \rangle, \langle z_0 \rangle$$
By (c), either $y_0 \in R \setminus F$ or $z_0 \in R \setminus F$ (or both). Without loss of generality, assume $x_1 = y_0 \in R \setminus F$. Therefore, 
$$x_0 = x_1 z_0 \implies x_0 \in \langle x_1 \rangle \implies
\langle x_0 \rangle \subsetneq \langle x_1 \rangle \quad (\mathrm{and}\ \langle x_1 \rangle \neq R).$$ 
Since $x_1$ not unit, we continue same argument on $x_1$ and get
$$\langle x_0 \rangle \subsetneq \langle x_1 \rangle \subsetneq \langle x_2 \rangle \subsetneq \cdots$$
contradicting (ACCP) which says there exists $N \in \mathbb{N}$ such that
$$\langle x_n \rangle = \langle x_{n+1} \rangle = \langle x_{n+2} \rangle = \cdots \quad $$
This is a contradiction, and consequently there is no such $x_0 \in R \setminus F$, In other words, $R = F$.
\end{proof}


\begin{theorem}
Every PID is a factorization domain.
\end{theorem}

\begin{proof}
Every PID satisfies the ACCP by the first lemma, and every ring satisfying the ACCP is a factorization domain by the second lemma.
\end{proof}

We have shown that in a PID, every nonzero nonunit factors into irreducibles. We now address uniqueness.

\begin{proposition}
Let $R$ be an integral domain and suppose
\[
 x_1 x_2 \cdots x_r \sim y_1 y_2 \cdots y_n,
\]
where $u \in U(R)$ is a unit, $x_1, \dots, x_r$ are primes, and $y_1, \dots, y_n$ are irreducible nonunits. Then $r = n$ and there exists a permutation $\sigma \in S_r$ such that $x_i \sim y_{\sigma(i)}$ for all $i$.
\end{proposition}

\begin{proof}
We proceed by induction on $r$.

\medskip
\noindent \textit{Base case} ($r = 0$): The left-hand side equals a unit, so $y_1 \cdots y_n \sim 1$. This forces each $y_j$ to be a unit, contradicting the assumption that $y_j$ are nonunits. Hence $n = 0$.

\medskip
\noindent \textit{Inductive step}: Assume the result holds for all products of fewer than $r$ primes. Consider 
$$u \cdot x_1 \cdots x_r = y_1 \cdots y_n.$$ Since $x_r$ is prime and $x_r \mid y_1 \cdots y_n$, we have $x_r \mid y_j$ for some $j$. Write $y_j = x_r \cdot a$ for some $a \in R$. Since $y_j$ is irreducible and $x_r$ is a nonunit, $a$ must be a unit, so $y_j \sim x_r$. Canceling $x_r$ from both sides (valid since $R$ is an integral domain) gives
\[
u \cdot x_1 \cdots x_{r-1} = y_1 \cdots \widehat{y_j} \cdots y_n \ \ 
\Longleftrightarrow \ \ x_1 \cdots x_{r-1} \sim y_1 \cdots \widehat{y_j} \cdots y_n
\]
where $\widehat{y_j}$ denotes that $y_j$ is omitted. By the inductive hypothesis, $r - 1 = n - 1$ (so $r = n$) and there exists $\sigma \in S_{r-1}$ matching $x_i \sim y_{\sigma(i)}$ for $i = 1, \dots, r-1$. Extending $\sigma$ by setting $\sigma(r) = j$ yields the desired permutation in $S_r$.
\end{proof}

\begin{corollary}
Let $R$ be a PID. Then every nonzero nonunit $x \in R$ can be written as a finite product of irreducibles (= primes), and this factorization is unique up to order and associates.
\end{corollary}

\noindent This generalizes the Fundamental Theorem of Arithmetic (the case $R = \mathbb{Z}$) to all PIDs.

\begin{definition}[Unique Factorization Domain]
Let $R$ be an integral domain. We say $R$ is a \textbf{unique factorization domain} (UFD) if every nonzero nonunit $x \in R$ can be written as a finite product of irreducibles, and this factorization is unique up to order and associates.
\end{definition}

The corollary above shows that every PID is a UFD. In a UFD, the notions of prime and irreducible also coincide.

\begin{proposition}
Let $R$ be a UFD. Then $p \in R$ is prime if and only if $p$ is irreducible.
\end{proposition}

\begin{proof}
Prime implies irreducible in any integral domain. Conversely, suppose $p$ is irreducible and $p \mid ab$, so $pc = ab$ for some $c \in R$. Write $a = a_1 \cdots a_s$, $b = b_1 \cdots b_t$, and $c = c_1 \cdots c_u$ as products of irreducibles. Then
\[
p \cdot c_1 \cdots c_u = a_1 \cdots a_s \cdot b_1 \cdots b_t.
\]
By uniqueness of factorization, $p$ must be an associate of some $a_i$ or some $b_j$. In the former case $p \mid a$, and in the latter $p \mid b$. Hence $p$ is prime.
\end{proof}

\noindent \textbf{Summary.} We have established the following strict chain of inclusions among classes of integral domains:
\[
\text{Fields} \subsetneq \text{PIDs} \subsetneq \text{UFDs} \subsetneq \text{Integral Domains.}
\]
Examples separating these classes: 
\begin{itemize}
    \item $\mathbb{Z}[\sqrt{-5}]$ is an integral domain but not a UFD, since irreducible does not imply prime (Homework), while irreducible and prime are the same for UFDs;
    \item $\mathbb{Z}[x]$ is a UFD but not a PID - we have seen it is not a PID, and will prove that it is a UFD in Section \ref{sec:polyrings} below.
\end{itemize}


\section{Euclidean Domain}

We now establish the connection between Euclidean domains and principal ideal domains.

\begin{definition}[Euclidean Function]
Let $R$ be an integral domain. A function 
\[
N: R \setminus \{0\} \to \mathbb{N} \cup \{0\}
\]
is called a \textbf{Euclidean function} on $R$ if for all $a, b \in R$ with $b \neq 0$, either 
\begin{itemize}
    \item[(i)] $b \mid a$; or
    \item[(ii)] there exist $q, r \in R$ with $r \neq 0$ such that $a = bq + r$ and $N(r) < N(b)$.
\end{itemize}
\end{definition}

\begin{example}
We have already seen the following examples of Euclidean functions:
\begin{itemize}
    \item $R = \mathbb{Z}$, $N(m) := |m|$. Then (ii) is the division algorithm for integers.
    \item $R = \mathbb{F}[x]$ (polynomials over a field), $N(p(x)) := \deg(p)$. Then (ii) is the division algorithm for polynomials.
\end{itemize}
\end{example}

\begin{definition}[Euclidean Domain]
An integral domain $R$ is a \textbf{Euclidean domain (ED)} if there exists at least one Euclidean function on $R$. 

If a Euclidean function $N$ satisfies $N(ab) = N(a)N(b)$ for all nonzero $a, b \in R$, we say $N$ is \textbf{multiplicative}.
\end{definition}

\begin{remark}
The Euclidean function $N(a) = |a|$ on $\mathbb{Z}$ is multiplicative. However, the degree function $N(p) = \deg(p)$ on $\mathbb{F}[x]$ is not multiplicative. To make it multiplicative, one can take $N'(p) := 2^{\deg(p)}$ which is still a (multiplicative) Euclidean function on $\mathbb{F}[x]$.
\end{remark}

\begin{theorem}
Let $R$ be an integral domain.
\begin{itemize}
    \item[(a)] If $R$ is a field, then $R$ is a Euclidean domain.
    \item[(b)] If $R$ is a Euclidean domain, then $R$ is a principal ideal domain.
\end{itemize}
\end{theorem}

\begin{proof}
\noindent \textit{(a):} Let $R$ be a field and define $N: R \setminus \{0\} \to \mathbb{N} \cup \{0\}$ by $N(a) := 1$ for all $a \in R \setminus \{0\}$. For any $a, b \in R$ with $b \neq 0$, since $R$ is a field, $b$ is a unit, so $b \mid a$. Thus condition (i) of the Euclidean function is satisfied, making $N$ a Euclidean function on $R$.

\medskip
\noindent \textit{(b):} Let $N$ be a Euclidean function on $R$, and let $I$ be a nonzero ideal of $R$. Choose $a \in I \setminus \{0\}$ such that $N(a)$ is minimal among all nonzero elements in $I$ (such an element exists since $N$ takes values in $\mathbb{N} \cup \{0\}$ and $I$ is nonzero). We claim that $I = \langle a \rangle$.

For any $i \in I$, apply the definition of Euclidean function to the pair $(i, a)$. Either $a \mid i$, or there exist $q, r \in R$ with $r \neq 0$ such that $i = aq + r$ with $N(r) < N(a)$.

In the first case, $i \in \langle a \rangle$.

In the second case, we have $r = i - aq$. Since $i, a \in I$ and $I$ is an ideal, we have $r \in I$. Moreover, $r \neq 0$ and $N(r) < N(a)$, contradicting the minimality of $N(a)$. Therefore, this case cannot occur.

Thus for all $i \in I$, we have $a \mid i$, which means $i \in \langle a \rangle$. Therefore $I \subseteq \langle a \rangle$. Since $a \in I$, we also have $\langle a \rangle \subseteq I$. Hence $I = \langle a \rangle$, so $R$ is a principal ideal domain.
\end{proof}

\noindent \textbf{Summary.} We have established the following strict chain of inclusions among classes of integral domains:
\[
\text{Fields} \subsetneq \text{EDs} \subsetneq \text{PIDs} \subsetneq \text{UFDs} \subsetneq \text{Integral Domains}.
\]
It is harder to get an example of a PID but not an ED:
\begin{example}[PID but not ED]
An example of a principal ideal domain that is not a Euclidean domain is more subtle. We present the ring
\[
R = \mathbb{Z}\left[\frac{1+\sqrt{-19}}{2}\right] = \left\{a + b\frac{1+\sqrt{-19}}{2} \ \Bigg|\ a, b \in \mathbb{Z}\right\}.
\]
The reason why $R$ is a PID is one of the main content of MAT 5210 -- Advanced Abstract Algebra I, based on the fact that the class number of the {\it number field} $\mathbb{Q}(\sqrt{-19})$ is $1$.

%(Note: The quadratic fields with discriminant $d < 0$ and class number $1$ are finite in number. The cases $d \in \{-1, -2, -3, -7, -11, -19, -43, -67, -163\}$ give class number $1$. The case $d = -19$ is one such example. A related interesting phenomenon is that $e^{\pi \sqrt {163}}$ is very close to an integer!)

As for why $R$ is not a Euclidean domain, it is left as an Homework exercise. 

%For $R$ to be a Euclidean domain, it would need to admit a Euclidean function $N: R \setminus \{0\} \to \mathbb{N} \cup \{0\}$.  However, it can be shown that no such function exists for $R = \mathbb{Z}\left[\frac{1+\sqrt{-19}}{2}\right]$ by analyzing the divisibility properties and the structure of the norm map. Specifically:
\end{example}


\begin{proposition}
The Gaussian integers $\mathbb{Z}[i] := \{a + bi \mid a, b \in \mathbb{Z}\}$ form a Euclidean domain with Euclidean function 
\[
N(a + bi) = a^2 + b^2.
\]
\end{proposition}

\begin{proof}
Let $\alpha, \beta \in \mathbb{Z}[i]$ with $\beta \neq 0$. Consider $\frac{\alpha}{\beta} = c_1 + c_2 i$ where $c_1, c_2 \in \mathbb{Q}$. 

Let $n_1, n_2 \in \mathbb{Z}$ be integers closest to $c_1$ and $c_2$ respectively, so that $|n_j - c_j| \leq \frac{1}{2}$ for $j = 1, 2$. Set $q := n_1 + n_2 i \in \mathbb{Z}[i]$.
Then 
\[
\frac{\alpha}{\beta} - q = (c_1 - n_1) + (c_2 - n_2)i.
\]
Multiplying by $\beta$ gives 
\[
\alpha - \beta q = [(c_1 - n_1) + (c_2 - n_2)i] \beta.
\]
Set $r := \alpha - \beta q$. If $r = 0$, then $\beta \mid \alpha$. Otherwise, $r \neq 0$ and 
\begin{align*}
N(r) &= N[(c_1 - n_1) + (c_2 - n_2)i] \cdot N(\beta) \\
&= [(c_1 - n_1)^2 + (c_2 - n_2)^2] \cdot N(\beta) \\
&\leq \left[\left(\frac{1}{2}\right)^2 + \left(\frac{1}{2}\right)^2\right] N(\beta) \\
&= \frac{1}{2} N(\beta) < N(\beta).
\end{align*}
Thus the Euclidean property holds, and $\mathbb{Z}[i]$ is a Euclidean domain.
\end{proof}


\noindent \textbf{Classification of Gaussian Primes.} Since $\mathbb{Z}[i]$ is a ED (and hence a PID), irreducible elements are precisely prime elements. 

We now classify the primes in $\mathbb{Z}[i]$. Before doing so, we give two preliminary lemmas:
\begin{lemma}
The units in $\mathbb{Z}[i]$ are $U(\mathbb{Z}[i]) = \{\pm 1, \pm i\}$. In other words, $\alpha \in \mathbb{Z}[i]$ is a unit if and only if $N(\alpha) = 1$.
\end{lemma}

\begin{lemma}
If $\beta \in \mathbb{Z}[i]$ satisfies $N(\beta) = p$ where $p$ is a prime integer, then $\beta$ is prime in $\mathbb{Z}[i]$.
\end{lemma}

\begin{proof}
Suppose $\beta = \gamma \delta$ for some $\gamma, \delta \in \mathbb{Z}[i]$. Then 
\[
p = N(\beta) = N(\gamma) N(\delta).
\]
Since $p$ is prime, either $N(\gamma) = 1$ or $N(\delta) = 1$. Thus $\gamma$ or $\delta$ is a unit, so $\beta$ is irreducible. Since $\mathbb{Z}[i]$ is a UFD, $\beta$ is prime.
\end{proof}

We divide into two cases to determine all primes in $\mathbb{Z}[i]$:

\medskip
\noindent{\underline{\bf Case 1: Elements of the form $a + 0i$ (or $0 + ai$).}} Since $a + 0i$ and $0 + ai$ only differ by a unit $i \in U(\mathbb{Z}[i])$, it suffices to check whether $a + 0i$ is prime or not. Obviously, if $a$ is a composite integer, then they are obviously not prime. So we assume $a = p$ is a prime integer. In such a case, one has:
\begin{lemma}
Let $p$ be a prime integer in $\mathbb{Z}$. Then either 
\begin{itemize}
\item[(i)] $p$ is prime in $\mathbb{Z}[i]$, or \item[(ii)] $p = (a + bi)(a - bi)$ where $a + bi$ and $a - bi$ are primes in $\mathbb{Z}[i]$ with $a, b \neq 0$.
\end{itemize}
\end{lemma}

\begin{proof}
Suppose $p$ is not prime in $\mathbb{Z}[i]$. Then $p = \gamma \delta$ for nonunits $\gamma, \delta \in \mathbb{Z}[i]$. Taking norms,
\[
p^2 = N(p) = N(\gamma) N(\delta).
\]
Thus $N(\gamma) = N(\delta) = p$. By the previous lemma, both $\gamma$ and $\delta$ are prime in $\mathbb{Z}[i]$.

Since $p \in \mathbb{Z}$ is real, $\bar{p} = p$. If $p = \gamma \delta$, then $p = \overline{\gamma \delta} = \bar{\gamma} \bar{\delta}$. Write $\gamma = a + bi$. Then $\bar{\gamma} = a - bi$, and by uniqueness of factorization in the UFD $\mathbb{Z}[i]$, we have $\bar{\gamma} \sim \delta$ (up to units and reordering). Since $N(\gamma) = N(\bar{\gamma}) = p$, we have $p = \gamma \bar{\gamma} = (a + bi)(a - bi)$.
\end{proof}


\begin{theorem}
Let $p$ be a prime integer. Then $p$ is prime in $\mathbb{Z}[i]$ if and only if $p \equiv 3 \pmod{4}$.
\end{theorem}

\begin{proof}
We prove the contrapositive in both directions.


($\Leftarrow$) Suppose $p$ is not prime in $\mathbb{Z}[i]$, then by the previous lemma, $p = a^2 + b^2$ for some integers $a, b$. Now $a^2 + b^2 \equiv 0, 1, \text{ or } 2 \pmod{4}$ depending on the parities of $a$ and $b$. Thus $p \not\equiv 3 \pmod{4}$.

($\Rightarrow$) Suppose $p \not\equiv 3 \pmod{4}$. There are two possibilities:
\begin{itemize}
    \item If $p = 2$, then $2 = (1+i)(1-i)$, so $2$ is not prime in $\mathbb{Z}[i]$.
    \item If $p$ is an odd prime with $p \equiv 1 \pmod{4}$, then by Lagrange's lemma (see Homework), 
    $$p \ |\ m^2 + 1$$
    (in $\mathbb{Z}$) for some integer $m$. So we have $p\ |\ (m+i)(m-i)$ in $\mathbb{Z}[i]$.

    We {\it claim} that $p \nmid (m+i)$ and $p \nmid (m-i)$ in $\mathbb{Z}[i]$: Suppose on contrary that $p \mid (m+i)$, then there exists $a + bi \in \mathbb{Z}[i]$ such that
    $$p(a+bi) = m+i.$$
    However, by comparing the imaginary parts on both sides of the equation, one has $pb = 1$, which is impossible. Similar argument also holds for $p \nmid (m-i)$. Consequently, $p$ is not prime in $\mathbb{Z}[i]$.
\end{itemize}
 
\end{proof}

\noindent{\underline{\bf Case 2: Elements of the form $a+bi \in \mathbb{Z}[i]$ with $a, b \neq 0$.} In such a case, one has:
\begin{theorem}
Let $a + bi \in \mathbb{Z}[i]$ with $a, b \neq 0$. Then $a + bi$ is prime in $\mathbb{Z}[i]$ if and only if $a^2 + b^2 = p$ for some prime integer $p$.
\end{theorem}

\begin{proof}
($\Leftarrow$) If $a^2 + b^2 = p$ is prime, then $N(a + bi) = p$, so $a + bi$ is prime.

($\Rightarrow$) Suppose $a + bi$ is prime in $\mathbb{Z}[i]$ with $a, b \neq 0$. Then $\overline{a + bi} = a - bi$ is also prime (Exercise). 

Consider the factorization of $(a + bi)(a - bi) = a^2 + b^2$ into primes in $\mathbb{Z}$. Write 
$$(a + bi)(a - bi) = a^2 + b^2 = p_1 \cdots p_r,$$ 
where each $p_i \in \mathbb{Z}$ is a prime integer. Note that the left hand side has only two irreducible factors in $\mathbb{Z}[i]$, hence by uniqueness of factorization, we have $r \leq 2$ or else there are at least three irreducible factors on the right hand side.

If $r = 2$, then $(a+bi)(a-bi) = p_1 p_2$. Then both $p_1$ and $p_2$ are irreducible (hence prime) in $\mathbb{Z}[i]$. By unique factorization, we have $a+bi \sim p_i$ for some $i$, meaning $a+bi = u \cdot p_i$ for a unit $u \in \{\pm 1, \pm i\}$. But then $a + bi = \pm p_i$ or $a + bi = \pm i p_i$, contradicting $a, b \neq 0$.

Therefore $r = 1$, so $a^2 + b^2 = p_1$ is prime in $\mathbb{Z}$.
\end{proof}

Combining the two cases, one has a full classification of prime elements in $\mathbb{Z}[i]$:
\begin{corollary}
The primes in $\mathbb{Z}[i]$ are precisely the elements of the following forms (and their associates):
\begin{itemize}
    \item[(i)] $p$ is prime integer satisfying $p \equiv 3 \pmod{4}$ (which remain prime);
    \item[(ii)] $a + bi$ with $a, b \neq 0$ and $a^2 + b^2 = p$, such that 
    \begin{itemize}
        \item $p = 2$, or
        \item $p$ is a prime integer satisfying $p \equiv 1 \pmod{4}$.
    \end{itemize} 
\end{itemize}
\end{corollary}





\begin{corollary}[Fermat's Theorem on Sums of Two Squares]
If $p$ is a prime integer with $p \equiv 1 \pmod{4}$, then there exist integers $a, b$ such that $p = a^2 + b^2$.
\end{corollary}

\begin{proof}
If $p \equiv 1 \pmod{4}$, then $p$ is not prime in $\mathbb{Z}[i]$ by the previous theorem. Therefore, 
$$p = (a+bi)(a-bi)$$ 
for some $a+bi$ prime in $\mathbb{Z}[i]$ with $a, b \neq 0$. So $p = a^2 + b^2$.
\end{proof}

\begin{example}
We have 
$$5 = 1^2 + 2^2,\quad 13 = 2^2 + 3^2,\quad 17 = 1^2 + 4^2,\quad 29 = 2^2 + 5^2,\quad \cdots$$
\end{example}

\section{Polynomial Rings} \label{sec:polyrings}

In this section, we study rings of polynomials $R[x]$. A main reason why polynomial rings are important is the following:

\begin{theorem}
Let $\mathbb{F}$ be a field and suppose $f(x) \in \mathbb{F}[x]$ is irreducible. Then $\mathbb{F}[x] / \langle f(x) \rangle$ is a field.
\end{theorem}

\begin{proof}
Since $\mathbb{F}[x]$ is a Euclidean Domain, it is a Principal Ideal Domain. The ideal $I = \langle f(x) \rangle$ is prime (= maximal) because $f(x)$ is irreducible (= prime). Therefore, $I = \langle f(x) \rangle$ is a maximal ideal, and the result follows from the general theory of maximal ideals.
\end{proof}

As a consequence, there is an injective homomorphism of fields 
$$\phi: \mathbb{F} \hookrightarrow \mathbb{F}[x] / \langle f(x) \rangle =: \mathbb{K}$$ 
defined by 
$$\phi(a) := a + \langle f(x) \rangle.$$ 
This allows us to ``extend'' $\mathbb{F}$ to a larger field.

In Galois Theory, one tries to understand roots of (irreducible) polynomials $f(x) \in \mathbb{F}[x]$. Let $\bar{x} := x + \langle f(x) \rangle \in \mathbb{K}$. Then in $\mathbb{K}$, we have 
\[
f(\bar{x}) = f(x) + \langle f(x) \rangle = 0_{K}.
\]
So $\bar{x}$ is a root of $f(x)$ in $\mathbb{K}$! So the above construction gives a way to construct roots of any (irreducible) polynomials.

\medskip
Under such perspective, it is important to determine when a polynomial is irreducible. In this section, we present methods for checking whether $f(x) \in \mathbb{Q}[x]$ (or $\mathbb{Z}[x]$) is irreducible. Although $\mathbb{Z}$ is \textbf{not} a field, it is useful for understanding irreducibility of polynomials $f(x) \in \mathbb{Q}[x]$ with integer coefficients.

\begin{lemma}
Let $\phi: R \to S$ be a unital ring homomorphism (e.g., $\mathbb{Z} \hookrightarrow \mathbb{Q}$ or $\mathbb{Z} \twoheadrightarrow \mathbb{Z}_p$). Then the map $\tilde{\phi}: R[x] \to S[x]$ defined by 
\[
\tilde{\phi}(a_0 + a_1 x + \cdots + a_n x^n) := \phi(a_0) + \phi(a_1) x + \cdots + \phi(a_n) x^n
\]
is a unital ring homomorphism.
\end{lemma}

\begin{theorem}[Reduction Modulo Prime Test]
Let $f \in \mathbb{Z}[x]$ be monic and let $p$ be a prime integer. If $\tilde{\phi}(f) \in \mathbb{Z}_p[x]$ is irreducible, then $f$ is irreducible in $\mathbb{Z}[x]$.
\end{theorem}

\begin{proof}
Suppose $f = gh$ for some $g, h \in \mathbb{Z}[x]$. We must show that $g$ or $h$ is a unit.

Applying $\tilde{\phi}$, we have 
$$\tilde{f} = \tilde{g} \cdot \tilde{h} \in \mathbb{Z}_p[x].$$ Since $f$ is monic, both $g$ and $h$ must be monic as well (up to units $\pm 1$). Therefore, $\tilde{f}$, $\tilde{g}$, and $\tilde{h}$ are all monic with 
\begin{center}
$\deg(\tilde{g}) = \deg(g)$ and $\deg(\tilde{h}) = \deg(h)$.
\end{center}
Since $\tilde{f}$ is irreducible, either $\tilde{g}$ or $\tilde{h}$ is a unit. In other words, $\deg(\tilde{g}) = 0$ or $\deg(\tilde{h}) = 0$. Therefore, either $\deg(g) = 0$ or $\deg(h) = 0$, which means $g$ or $h$ is a constant polynomial. Since $f$ is monic with degree greater than $0$, this constant must be $\pm 1$, making $g$ or $h$ a unit in $\mathbb{Z}[x]$. Thus $f$ is irreducible.
\end{proof}

\begin{example}
Consider $f(x) = x^3 + 3x + 7$ and reduce modulo $p = 2$. We get 
$$\tilde{f}(x) = x^3 + x + 1 \in \mathbb{Z}_2[x].$$
To show $\tilde{f}$ is irreducible in $\mathbb{Z}_2[x]$, note that if $\tilde{f}$ (or more generally any cubic polynomial) factors, then it must have a linear factor. Hence by the factor theorem, either $\tilde{f}(0) = 0$ or $\tilde{f}(1) = 0$ in $\mathbb{Z}_2$. However, we compute:
\[
\tilde{f}(0) = 1 \neq 0 \pmod{2}, \quad \tilde{f}(1) = 1 + 1 + 1 = 1 \neq 0 \pmod{2}.
\]
Therefore, $\tilde{f}$ has no linear factors in $\mathbb{Z}_2[x]$, so it is irreducible. Thus $f(x) = x^3 + 3x + 7$ is irreducible in $\mathbb{Z}[x]$.
\end{example}

\begin{definition}[Primitive Polynomial]
A polynomial $f(x) = a_n x^n + \cdots + a_0 \in \mathbb{Z}[x]$ is called \textbf{primitive} if $\gcd(a_n, \ldots, a_0) = 1$.
\end{definition}


\begin{theorem}[Eisenstein's Criterion]
Let $f(x) = a_n x^n + \cdots + a_1 x + a_0 \in \mathbb{Z}[x]$ be a primitive polynomial. Suppose $p$ is a prime integer such that:
\begin{enumerate}
    \item[(1)] $p \mid a_i$ for all $0 \le i < n$;
    \item[(2)] $p \nmid a_n$;
    \item[(3)] $p^2 \nmid a_0$.
\end{enumerate}
Then $f(x)$ is irreducible in $\mathbb{Z}[x]$.
\end{theorem}

\begin{proof}
Suppose $f = gh$ for some $g, h \in \mathbb{Z}[x]$. We must show that $g$ or $h$ is a unit. Consider the reduction map $\tilde{\phi}: \mathbb{Z}[x] \to \mathbb{Z}_p[x]$. By condition (1), we have 
\[
\tilde{f}(x) = \overline{a_n} x^n
\]
where $\overline{a_n} \neq 0$ in $\mathbb{Z}_p$ by condition (2).

From $\tilde{f} = \tilde{g} \cdot \tilde{h}$ and $\tilde{f} \sim x^n$ in $\mathbb{Z}_p[x]$ (since $\overline{a_n}$ is a unit in $\mathbb{Z}_p$), we must have 
\begin{center}
$\tilde{g} \sim x^i$ and $\tilde{h} \sim x^{n-i}$
\end{center}
for some $0 \le i \le n$ by unique factorization of $\mathbb{Z}_p[x]$. Suppose on contrary $0 < i < n$. Then the constant term of $\tilde{g}$ is $0$ in $\mathbb{Z}_p$, which means $p$ divides the constant term of $g$. Similarly, $p$ divides the constant term of $h$. Therefore, $p^2$ divides the constant term of $gh = f$, contradicting condition (3).

Thus either $i = 0$ or $i = n$, which means either $\tilde{g}$ or $\tilde{h}$ is a constant in $\mathbb{Z}_p[x]$. Finally, Ssnce 
$$\deg(g) + \deg(h) = \deg(f) = \deg(\tilde{f}) = \deg(\tilde{g}) + \deg(\tilde{h})$$ 
and 
\begin{center}
$\deg(\tilde{g}) \le \deg(g)$, $\deg(\tilde{h}) \le \deg(h)$, 
\end{center}
we have $\deg(\tilde{g}) = \deg(g)$ and $\deg(\tilde{h}) = \deg(h)$.

Therefore, either $\deg(g) = 0$ or $\deg(h) = 0$, so $g$ or $h$ is a constant polynomial. Since $f$ is primitive, this constant must be $\pm 1$, making $g$ or $h$ a unit. Thus $f$ is irreducible.
\end{proof}

\begin{example}
Consider $g(x) = x^4 + 1$. Define $f(x) = g(x+1) = (x+1)^4 + 1$. Expanding:
\[
f(x) = x^4 + 4x^3 + 6x^2 + 4x + 2.
\]
Taking $p = 2$, we verify Eisenstein's criterion:
\begin{enumerate}
    \item $2 \mid 4, 6, 4, 2$ (all non-leading coefficients);
    \item $2 \nmid 1$ (the leading coefficient);
    \item $2^2 = 4 \nmid 2$ (the constant term).
\end{enumerate}
Therefore, $f(x)$ is irreducible in $\mathbb{Z}[x]$, and hence $g(x) = x^4 + 1$ is also irreducible in $\mathbb{Z}[x]$.
\end{example}

Now we address irreducibility in $\mathbb{Q}[x]$.

\begin{theorem}[Gauss' Lemma]
Let $f \in \mathbb{Z}[x] (\subseteq \mathbb{Q}[x])$ be a nonconstant polynomial with integer coefficients. Then 
\begin{center}
$f(x)$ is irreducible in $\mathbb{Z}[x]$ \quad  $\Longleftrightarrow$ \quad $f(x)$ is primitive, and is irreducible in $\mathbb{Q}[x]$.
\end{center}
\end{theorem}

\begin{proof}
$(\Leftarrow)$ Assume $f \in \mathbb{Z}[x]$ is primitive, irreducible in $\mathbb{Q}[x]$. Suppose 
$$f = gh, \quad \quad g, h \in \mathbb{Z}[x]$$ 
We must show that $g$ or $h$ is a unit in $\mathbb{Z}[x]$.

Since $g, h \in \mathbb{Z}[x] \subseteq \mathbb{Q}[x]$ and $f$ is irreducible in $\mathbb{Q}[x]$, either $g$ or $h$ must be a unit in $\mathbb{Q}[x]$, i.e., a nonzero constant. Say $g = c \in \mathbb{Q}$. But $g \in \mathbb{Z}[x]$, we have $c \in \mathbb{Z}$. Moreover, since $f$ is primitive, $\gcd(\text{coefficients of } f) = 1$. If $c \neq \pm 1$, then $c$ divides all coefficients of $f = gh = c \cdot h$, contradicting primitivity of $f$. Therefore, $c = \pm 1$, so $g$ is a unit in $\mathbb{Z}[x]$.

$(\Rightarrow)$ Suppose $f$ is irreducible in $\mathbb{Z}[x]$.
Then $f$ must be primitive, otherwise it has a nonunit constant factor $d := \gcd(\text{coefficients of } f)$. We can write 
$$f = d \cdot f'$$ 
where $f' \in \mathbb{Z}[x]$ and $\deg(f') = \deg(f) > 0$. Since neither $d$ nor $f'$ is a unit in $\mathbb{Z}[x]$, this contradicts irreducibility of $f$. Thus $f$ is primitive.

Next, we show $f$ is irreducible in $\mathbb{Q}[x]$. Suppose 
$$f = gh \quad \quad g, h \in \mathbb{Q}[x].$$ 
We must show that $g$ or $h$ is a unit in $\mathbb{Q}[x]$.
Let \(\lambda \in \mathbb{N}\) be the smallest positive integer such that there exists \( \alpha \in \mathbb{Q}\setminus\{0\} \) with
\[
h' := \alpha h \in \mathbb{Z}[x]
\qquad \text{and} \qquad
g' := \lambda \alpha^{-1} g \in \mathbb{Z}[x].
\]
Then \(
g'h' = (\lambda \alpha^{-1} g)(\alpha h)=\lambda gh=\lambda f.
\) in $\mathbb{Z}[x]$.

\noindent \textbf{Claim:} \(\lambda = 1\).

\noindent \textit{Proof of claim:} Suppose for contradiction that \(\lambda > 1\). Then there exists a prime \(p\) such that \(p \mid \lambda\). Since
\[
\lambda f = g'h',
\]
all coefficients of \(g'h'\) are divisible by \(p\). Hence, reducing modulo \(p\), we get
\[
\widetilde{g'}\widetilde{h'} = 0
\quad \text{in }\ \mathbb{Z}_p[x].
\]
Since \(\mathbb{Z}_p[x]\) is an integral domain, it follows that
\(
\widetilde{g'}=0\) or \(\widetilde{h'}=0.\)
Equivalently,
\[
p \mid g' \qquad \text{or} \qquad p \mid h'
\qquad \text{in } \mathbb{Z}[x].
\]
If \(p \mid g'\), then
\[
\frac{1}{p}g'=\frac{\lambda}{p}\alpha^{-1}g \in \mathbb{Z}[x].
\]
Together with \(\alpha h \in \mathbb{Z}[x]\), this contradicts the minimality of \(\lambda\).

If \(p \mid h'\), then
\[
\frac{1}{p}h'=\frac{\alpha}{p}h \in \mathbb{Z}[x].
\]
Also,
\[
\frac{\lambda}{p}\left(\frac{\alpha}{p}\right)^{-1}g
=
\frac{\lambda}{p}\cdot \frac{p}{\alpha}g
=
\lambda \alpha^{-1}g
\in \mathbb{Z}[x].
\]
Thus, with \(\alpha' := \alpha/p\), we have
\[
\alpha' h \in \mathbb{Z}[x]
\qquad \text{and} \qquad
\frac{\lambda}{p}(\alpha')^{-1}g \in \mathbb{Z}[x],
\]
again contradicting the minimality of \(\lambda\).

Therefore no prime divides \(\lambda\), so \(\lambda=1\). This proves the claim.

\medskip

Hence there exists \(\alpha \in \mathbb{Q}\setminus\{0\}\) such that
\[
\alpha^{-1}g \in \mathbb{Z}[x]
\qquad \text{and} \qquad
\alpha h \in \mathbb{Z}[x].
\]
Thus
\[
f = (\alpha^{-1}g)(\alpha h)
\]
is a factorization of \(f\) in \(\mathbb{Z}[x]\). Since \(f\) is irreducible in \(\mathbb{Z}[x]\), either \(\alpha^{-1}g\) or \(\alpha h\) is a unit in \(\mathbb{Z}[x]\). Hence either \(\alpha^{-1}g=\pm 1\) or \(\alpha h=\pm 1\). In the first case,
\(
g=\pm \alpha \in \mathbb{Q} \backslash \{0\},
\)
so \(g\) is a unit in \(\mathbb{Q}[x]\). In the second case,
\(
h=\pm \alpha^{-1}  \in \mathbb{Q} \backslash \{0\},
\) so \(h\) is a unit in \(\mathbb{Q}[x]\). Therefore \(f\) is irreducible in \(\mathbb{Q}[x]\).
\end{proof}

{\color{blue} We present a flawed proof here - it will be a good exercise to (a) find a counterexample of the proof below; and (b) which part of the proof goes wrong:

Next, we show $f$ is irreducible in $\mathbb{Q}[x]$. Suppose 
$$f = gh \quad \quad g, h \in \mathbb{Q}[x].$$ 
We must show that $g$ or $h$ is a unit in $\mathbb{Q}[x]$.
To do so, let $\lambda, \mu \in \mathbb{N}$ be the smallest positive integers such that $g' := \lambda g \in \mathbb{Z}[x]$ and $h' := \mu h \in \mathbb{Z}[x]$ have integer coefficients. Then 
\[
f = gh = \frac{g'}{\lambda} \cdot \frac{h'}{\mu},
\]
so that $\lambda \mu f = g' h'$ in $\mathbb{Z}[x]$.

\noindent \textbf{Claim:} $\lambda = 1$.

\noindent \textit{Proof of claim:} Suppose for contradiction that $\lambda > 1$, so there exists a prime $p \mid \lambda$. Then $p \mid \lambda \mu f$, so all coefficients of $\lambda \mu f \in \mathbb{Z}[x]$ are divisible by $p$. In particular, $p \mid g' h'$ in $\mathbb{Z}[x]$.

Consider the reduction map $\tilde{\phi}: \mathbb{Z}[x] \to \mathbb{Z}_p[x]$. We have $\tilde{g'} \cdot \tilde{h'} \equiv 0$ in $\mathbb{Z}_p[x]$. Since $\mathbb{Z}_p[x]$ is an integral domain, either $\tilde{g'} \equiv 0$ or $\tilde{h'} \equiv 0$ in $\mathbb{Z}_p[x]$. This means either $p \mid g'$ or $p \mid h'$.

If $p \mid g'$, then all coefficients of $g' = \lambda g$ are divisible by $p$. But then $(\lambda/p) g$ has integer coefficients and is a scalar multiple of $g$, contradicting the minimality of $\lambda$. Similarly, if $p \mid h'$, then $(\mu/p) h$ has integer coefficients, contradicting the minimality of $\mu$.

Therefore, $\lambda = 1$, and $g' = g \in \mathbb{Z}[x]$ and $h' = \mu h \in \mathbb{Z}[x]$. By the same argument applied to $\mu$, we also have $\mu = 1$, so $h \in \mathbb{Z}[x]$.

Now $f = gh$ with $g, h \in \mathbb{Z}[x]$. Since $f$ is irreducible in $\mathbb{Z}[x]$, either $g$ or $h$ is a unit in $\mathbb{Z}[x]$, which means either $g$ or $h$ is $\pm 1 \in \mathbb{Q}[x]$. Thus $f$ is irreducible in $\mathbb{Q}[x]$.
}

\begin{corollary}
$\mathbb{Z}[x]$ is a Unique Factorization Domain (but not a Principal Ideal Domain).
\end{corollary}

\begin{proof}
First, we show that every primitive polynomial $f \in \mathbb{Z}[x]$ with $\deg(f) \ge 1$ can be factored into primitive irreducible polynomials. We proceed by induction on $\deg(f)$.

\noindent \textit{Base case:} If $\deg(f) = 1$, then $f$ is irreducible.

\noindent \textit{Inductive step:} Assume the claim holds for all primitive polynomials of degree less than $k$. Let $f$ be primitive with $\deg(f) = k > 1$. Either:
\begin{itemize}
    \item $f$ is irreducible; or
    \item $f = h_1 h_2$ for some $h_1, h_2 \in \mathbb{Z}[x]$ with $0 < \deg(h_i) < k$.
\end{itemize}
In the second case, since $f$ is primitive, $h_1$ and $h_2$ must also be primitive (otherwise their gcd would divide $f$'s coefficients, contradicting primitivity). By induction, both $h_1$ and $h_2$ factor into primitive irreducibles, so $f$ does as well.

Now, to show factorization is unique, suppose 
$$f = g_1 \cdots g_n = h_1 \cdots h_m$$ 
where $g_i, h_j \in \mathbb{Z}[x]$ are primitive and irreducible. 

By Gauss' Lemma, $g_i$ and $h_j$ are also irreducible in $\mathbb{Q}[x]$. Since $\mathbb{Q}[x]$ is a principal ideal domain (hence a UFD), we have $n = m$ and there exists a permutation $\sigma \in S_n$ such that $g_i$ and $h_{\sigma(i)}$ are associates in $\mathbb{Q}[x]$. That is, $g_i = c_i h_{\sigma(i)}$ for some $c_i \in \mathbb{Q} \backslash \{0\}$.

Since $g_i, h_{\sigma(i)} \in \mathbb{Z}[x]$ and both are primitive, we must have $c_i \in \{\pm 1\}$. Therefore, $g_i \sim h_{\sigma(i)}$ in $\mathbb{Z}[x]$. Thus factorization into primitive irreducibles is unique up to order and associates.

For a general polynomial $f \in \mathbb{Z}[x]$, one can pull out the greatest common factor $d \in \mathbb{Z}$ of the coefficients of $f$, and write 
$f = d \cdot f'$
where $d \in \mathbb{Z}$ and $f'$ is primitive. One can factor $d$ uniquely in $\mathbb{Z}$, and factor $f'$ uniquely into primitive irreducibles in $\mathbb{Z}[x]$. The factorization of $f$ is therefore unique, making $\mathbb{Z}[x]$ a UFD.
\end{proof}

\begin{remark}
More generally, if $R$ is a unique factorization domain, then $R[x]$ is also a unique factorization domain. The proof uses the field of fractions $\mathbb{F} = \text{Frac}(R)$ and the fact that $\mathbb{F}[x]$ is a principal ideal domain (hence a UFD).
\end{remark}